---
title: "Beta diversity analysis with Bracken Xu et al chinese cohort data with read threshold 1 and kmer minimizer threshold 3"
author: "Tanya Karagiannis"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    theme: united
    toc: yes
    toc_collapse: no
    toc_float: yes
  pdf_document:
    toc: yes
---

```{r setup, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
library(dplyr)
library(ggplot2)
library(ggpubr)
library(ape)
library(DESeq2)
library(stringr)
library(reshape2)
library(graphics)

library(codebook)
library(future)
#library(vtable)
library(useful)

library(vegan)
library(data.table)

# Data Cleaning
library(janitor)

# EDA
library(skimr)
library(DataExplorer)

# ggplot2 Helpers
library(ggplot2)
library(ggsignif)
library(scales)
library(phyloseq)
library(tidyverse)
library(tidymodels)
#heatmap
library(pheatmap)
library(pheatmap)
library(ComplexHeatmap)
library(circlize)
library(dendextend)
library(dynamicTreeCut)
#microbiome and single cell specific packages
library(vegan)
library(mia)
library(scater)
#write excel sheets
library(writexl)


#specify file path
flpath <- "/restricted/projectnb/uh2-sebas/data/metagenomics/external_cohorts/Han_Chinese_Centenarian/processed_data/data_library/"
outpath <- "/restricted/projectnb/uh2-sebas/analysis/metagenomics/parallel_systems_comparison/02_beta_diversity_analysis/Chinese_cohort/output/"
meta.dir <- "/restricted/projectnb/necs/Data_library/Metabolomics/resources/Xu_plasma_metabolite_analysis/"
sra.dir <- "/restricted/projectnb/uh2-sebas/data/metagenomics/external_cohorts/Han_Chinese_Centenarian/metadata/"

otufl <- "bracken_renorm_chinese_phyloseq.R1.K3.10.22.2024.rds"
phenofl <- "phenotype.xlsx"

#set seed
set.seed(157)
```

This script is for diversity analysis using data after QC.

```{r user-defined functions}
## outlier function
is_outlier <- function(x) {
  return(x < quantile(x, 0.25) - 1.5 * IQR(x) | x > quantile(x, 0.75) + 1.5 * IQR(x))
}


# calculate relative abundance/normalized abundance
get_relabund <- function(indat, varname){
  relabund <- as.tibble(indat) %>%
    janitor::adorn_percentages("col") %>%
    column_to_rownames(var = varname) 
}
# 
source("/restricted/projectnb/uh2-sebas/analysis/metagenomics/ye_analyses/Scripts/UserDefinedFunctions/CTDS_function.R")

#function to aggregate otu abundances, calculate proportions and calculated CTDS based on taxonomic hierarchy
otu_ctds <- function(taxadata = biom_chinese_kraken.taxa, 
                     otudata=biom_chinese_kraken.otu, rank=rank, 
                     sample.meta=sample.meta){
  #load ctds function
  source("/restricted/projectnb/uh2-sebas/analysis/metagenomics/ye_analyses/Scripts/UserDefinedFunctions/CTDS_function.R")
  
  #prepare taxonomic and otu data to run diversity analysis
  #convert taxonomic table to a tibble with rownames as a column called tax.ID
  #filter taxonomic table to remove taxa with no species label
  biom_taxdata <- as_tibble(taxadata, rownames = "tax.ID") 
 # message(paste("taxa table rows:",nrow(biom_taxdata)))
  #convert otu table to a tibble with rownames as a column called tax.ID
  biom_otudata <- as_tibble(otudata, rownames = "tax.ID")
  
  #save sample names
  biom_samplenames <- colnames(otudata)
  #message(paste("taxa table rows:",nrow(biom_otudata)))

  
  #taxa dictionary: taxonomy table with taxa ID and taxa names
  taxa_dict <- biom_taxdata %>% 
    dplyr::select(tax.ID, {{rank}}) %>%
    unite("tax.name", {{rank}}, remove = FALSE)

  #use an inner_join function from dplyr to join data from both table based on common tax ID
  #create new variable taxanames to create unique
  biom_join <- dplyr::inner_join(taxa_dict, biom_otudata, by = "tax.ID")
  #message(paste("joined table rows:", nrow(biom_join))) 


  #remove rankings and transpose data for samples in rows and taxa IDs in columns
  #make sure the samples are set as rownames
  # biom_prop <- biom_join %>% 
  #   group_by(tax.name) %>% #use rank groupings
  #   summarise_if(is.numeric, sum) %>% #aggregate OTU abundances for each taxa in the rank
  #   select(c(tax.name, contains(".aggregrated.report"))) %>% #select rank and sample OTU abundances
  #   dplyr::rename(tax.ID = tax.name) %>%
  #   tibble::column_to_rownames(var = "tax.ID") %>% #make tax.ID variable to rownames
  #   t()
  
  biom_prop <- biom_join %>% 
    group_by(tax.name) %>% #use rank groupings
    summarise_if(is.numeric, sum) %>% #aggregate OTU abundances for each taxa in the rank
    select(c(tax.name, biom_samplenames)) %>% #select rank and sample OTU abundances
    column_to_rownames(var = "tax.name") %>% #make tax.name variable to rownames
    t() %>% #transpose data to have taxa in columns and samples as rows
    as_tibble(rownames = "sample.ID") %>% #make transposed data into a tibble with rownames as sample.ID variable
    janitor::adorn_percentages() %>% #use janitor package to get row proportions for each sample
    column_to_rownames(var = "sample.ID") #add sample IDs back to rownames
  

  #ctds function
  biom.res <- CTDS.score(biom_prop, metadata = sample.meta)
  return(biom.res)
  

}


prop_mat <- function(taxadata = biom_chinese_kraken.taxa, 
                     otudata=biom_chinese_kraken.otu, rank=rank, 
                     sample.meta=sample.meta){

  #convert taxonomic table to a tibble with rownames as a column called tax.ID
  #filter taxonomic table to remove taxa with no taxa classification in any level
  biom_taxdata <- as_tibble(taxadata, rownames = "tax.ID") %>% 
    unite("tax.name", kingdom:{{rank}}, remove = FALSE)

  #row number for taxa data after removing incomplete classification
  message(paste("taxa table rows:",nrow(biom_taxdata)))

  #convert otu table to a tibble with rownames as a column called tax.ID
  biom_otudata <- as_tibble(otudata, rownames = "tax.ID")
  #row number for otu data before removing incomplete classification
  message(paste("otu table rows:",nrow(biom_otudata)))

  #use an inner_join function from dplyr to join data from both table based on common tax ID
  biom_join <- inner_join(biom_taxdata, biom_otudata, by = "tax.ID")
  #row number for joined data
  message(paste("joined table rows:", nrow(biom_join))) #row number 11039 for the common taxa
  
  biom_taxid <- biom_join %>% select(c(tax.ID, tax.name))
  

  #remove rankings and transpose data for samples in rows and taxa IDs in columns
  #make sure the samples are set as rownames
  biom_prop <- biom_join %>% 
    group_by(tax.name) %>%
    summarise_if(is.numeric, sum) %>% 
    select(c(tax.name, contains(".aggregrated.report"))) %>%
    column_to_rownames(var = "tax.name") %>% #make tax.name variable to rownames
    t() %>% #transpose data to have taxa in columns and samples as rows
    as_tibble(rownames = "sample.ID") %>% #make transposed data into a tibble with rownames as sample.ID variable
    janitor::adorn_percentages() %>% #use janitor package to get row proportions for each sample
    column_to_rownames(var = "sample.ID") #add sample IDs back to rownames
  dim(biom_prop)
  useful::corner(biom_prop) #use this to visualize a corner of the data

  #rerun abundance table
  biom.res <- CTDS.score(biom_prop, metadata = sample.meta)
  return(biom.res)

}
```

# Read in phenotypic data


```{r, message=FALSE}
# Extract phenotypic data
#load clinical metadata information
#clinical.data <- read_excel(paste0(meta.dir, "DifferentialAbundance_Supplement.xlsx"), sheet = "Table S1 Metadata", skip = 1)
clinical.data <- read_excel(paste0(meta.dir, phenofl))

#add ethnicity variable
clinical.data$ethnicity <- "han_chinese"


#extract sample IDs and create family ID and group ID
sample.ID <- clinical.data %>% dplyr::select(1) %>% pull()

fam.group <- str_remove(sample.ID, "QCS-") %>%                       
              str_split("-") %>%
                unlist() 

fam.group <- matrix(fam.group, nrow=2)

#create family ID
fam.id <- fam.group[1,]

#create group ID
group.id <- fam.group[2,]
group.id[group.id == "1"] <- "Advanced Age"
group.id[group.id == "2" | group.id == "3"] <- "Offspring"


#add to clinical data
clinical.clean <- clinical.data %>%
  mutate(sample.ID = sample.ID, fam.id = fam.id, group = group.id) 
clinical.clean$"C-age (years)" <- as.integer(clinical.clean$"C-age (years)")


#load sample Run names and sample SRA names
SRA.data <- read_csv(paste0(sra.dir,"SraRunInfo.csv"))

SRA.clean <- SRA.data %>% 
  dplyr::select(Run, SampleName) 

#sample ID and sample SRA names
sample.dict <- read_excel(paste0(sra.dir, "sample name-sequencing ID sheet.xlsx"))

#combine files
sample.info <- dplyr::inner_join(SRA.clean, sample.dict, by = c("SampleName"))

pheno.data <- dplyr::inner_join(sample.info, clinical.clean, by = c("sample.ID")) %>% 
                                  dplyr::select(-c(sample.ID)) %>%
                                  mutate(stool_kit = Run, Sample=Run) %>%
                                  dplyr::rename(c_age_years="C-age (years)")
pheno.data$stool_kit <- as.character(pheno.data$stool_kit)
pheno.data$Sample <- as.character(pheno.data$Sample)


SAMPLE_phy <- sample_data(pheno.data %>% column_to_rownames(var="Sample"))
```


# phyloseq object
```{r read in data}
#load phyloseq obj
filtered_chinese_biom <- readRDS(paste0(flpath, otufl))

#add phenotypic data to phyloseq object
filtered_chinese_biom@sam_data <- SAMPLE_phy

#keep samples with age label
subset_chinese_biom <- subset_samples(filtered_chinese_biom, !is.na(c_age_years))

OTU_filtered <- subset_chinese_biom@otu_table@.Data
TAX_filtered <- subset_chinese_biom@tax_table@.Data

sample.meta <- sample_data(subset_chinese_biom) %>% 
  as_tibble(rownames="sample.ID")

#save sample data for permanova
samp.dat <- as.matrix(subset_chinese_biom@sam_data)
```



# Beta diversity across taxonomic hierarchy

## Create Tree Summarized Experiment object from Phyloseq object 

```{r}
#convert phyloseq object to a tree summarized experiment object
kraken2uniq_tse <- mia::makeTreeSummarizedExperimentFromPhyloseq(subset_chinese_biom)
kraken2uniq_tse


#agglomerate based on each taxonomic rank and save as alternative assays in the object
altExps(kraken2uniq_tse) <- mia::splitByRanks(kraken2uniq_tse)
kraken2uniq_tse


```

```{r}
#taxa diversity for species
species_tse <- altExp(kraken2uniq_tse,"species")
beta.div.species <- rbiom::beta.div(assay(species_tse, "counts"), method="bray-curtis", weighted=TRUE)
uniqueness.species <- apply(as.matrix(beta.div.species), 1, FUN = function(x) {min(x[x > 0])})

#join with metadata
b.species <- uniqueness.species %>% as_tibble(rownames="sample.ID") %>%
  dplyr::rename("uniqueness"=value) %>%
  dplyr::inner_join(sample.meta, by = "sample.ID")

#taxa diversity for genus
genus_tse <- altExp(kraken2uniq_tse,"genus")
beta.div.genus <- rbiom::beta.div(assay(genus_tse, "counts"), method="bray-curtis", weighted=TRUE)
uniqueness.genus <- apply(as.matrix(beta.div.genus), 1, FUN = function(x) {min(x[x > 0])})

#join with metadata
b.genus <- uniqueness.genus %>% as_tibble(rownames="sample.ID") %>%
  dplyr::rename("uniqueness"=value) %>%
  dplyr::inner_join(sample.meta, by = "sample.ID")

#taxa diversity for family
family_tse <- altExp(kraken2uniq_tse,"family")
beta.div.family <- rbiom::beta.div(assay(family_tse, "counts"), method="bray-curtis", weighted=TRUE)
uniqueness.family <- apply(as.matrix(beta.div.family), 1, FUN = function(x) {min(x[x > 0])})

#join with metadata
b.family <- uniqueness.family %>% as_tibble(rownames="sample.ID") %>%
  dplyr::rename("uniqueness"=value) %>%
  dplyr::inner_join(sample.meta, by = "sample.ID")

#taxa diversity for order
order_tse <- altExp(kraken2uniq_tse,"order")
beta.div.order <- rbiom::beta.div(assay(order_tse, "counts"), method="bray-curtis", weighted=TRUE)
uniqueness.order <- apply(as.matrix(beta.div.order), 1, FUN = function(x) {min(x[x > 0])})

#join with metadata
b.order <- uniqueness.order %>% as_tibble(rownames="sample.ID") %>%
  dplyr::rename("uniqueness"=value) %>%
  dplyr::inner_join(sample.meta, by = "sample.ID")

#taxa diversity for class
class_tse <- altExp(kraken2uniq_tse,"class")
beta.div.class <- rbiom::beta.div(assay(class_tse, "counts"), method="bray-curtis", weighted=TRUE)
uniqueness.class <- apply(as.matrix(beta.div.class), 1, FUN = function(x) {min(x[x > 0])})

#join with metadata
b.class <- uniqueness.class %>% as_tibble(rownames="sample.ID") %>%
  dplyr::rename("uniqueness"=value) %>%
  dplyr::inner_join(sample.meta, by = "sample.ID")

#taxa diversity for phylum
phylum_tse <- altExp(kraken2uniq_tse,"phylum")
beta.div.phylum <- rbiom::beta.div(assay(phylum_tse, "counts"), method="bray-curtis", weighted=TRUE)
uniqueness.phylum <- apply(as.matrix(beta.div.phylum), 1, FUN = function(x) {min(x[x > 0])})

#join with metadata
b.phylum <- uniqueness.phylum %>% as_tibble(rownames="sample.ID") %>%
  dplyr::rename("uniqueness"=value) %>%
  dplyr::inner_join(sample.meta, by = "sample.ID")
```

## Save bray-curtis distances

```{r}
#save pair-wise distances information to excel file
BC_list <- list(
  BC.phylum = as_tibble(as.matrix(beta.div.phylum), rownames="sample.ID"),
  BC.class = as_tibble(as.matrix(beta.div.class), rownames="sample.ID"),
  BC.order = as_tibble(as.matrix(beta.div.order), rownames="sample.ID"),
  BC.family = as_tibble(as.matrix(beta.div.family), rownames="sample.ID"),
  BC.genus = as_tibble(as.matrix(beta.div.genus), rownames="sample.ID"),
  BC.species = as_tibble(as.matrix(beta.div.species), rownames="sample.ID")
  )

write_xlsx(BC_list, paste0(outpath,"BC_bracken_chinese.R1.K3.xlsx"))
```

# Uniqueness across taxonomic levels

## Age (continuous)

```{r, fig.dim = c(20,10)}
p.s_age <- ggplot(b.species, aes(x=c_age_years, y=uniqueness)) + 
    geom_point(fill="blue") + 
    ylim(0,1) + 
    theme_bw(base_size = 20) +
    theme(axis.title.y=element_blank(), axis.title.x=element_blank(),
          axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
          plot.title = element_text(hjust = 0.5),
          legend.position = "none") +
    ggtitle("Species") 
    

#For genus level
p.g_age <- ggplot(b.genus, aes(x=c_age_years, y=uniqueness)) + 
    geom_point(fill="blue") + 
    ylim(0,1) +
    theme_bw(base_size = 20) +
    theme(axis.title.y=element_blank(), axis.title.x=element_blank(),
          axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
          plot.title = element_text(hjust = 0.5),
          legend.position = "none") +
    ggtitle("Genus") 
    

#For family level
p.f_age <- ggplot(b.family, aes(x=c_age_years, y=uniqueness)) + 
    geom_point(fill="blue") + 
    ylim(0,1) + 
    theme_bw(base_size = 20) +
    theme(axis.title.y=element_blank(), axis.title.x=element_blank(),
          axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
          plot.title = element_text(hjust = 0.5),
          legend.position = "none") +
  ggtitle("Family") 
    


#For order level
p.o_age <- ggplot(b.order, aes(x=c_age_years, y=uniqueness)) + 
    geom_point(fill="blue") + 
    ylim(0,1) + 
    theme_bw(base_size = 20) +
    theme(axis.title.y=element_blank(), axis.title.x=element_blank(),
          axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
          plot.title = element_text(hjust = 0.5),
          legend.position = "none") +
    ggtitle("Order")


#For class level
p.c_age <- ggplot(b.class, aes(x=c_age_years, y=uniqueness)) + 
    geom_point(fill="blue") + 
    ylim(0,1) + 
    theme_bw(base_size = 20) +
    theme(axis.title.y=element_blank(), axis.title.x=element_blank(),
          axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
          plot.title = element_text(hjust = 0.5),
          legend.position = "none") +
    ggtitle("Class")


#For phylum level
p.p_age <- ggplot(b.phylum, aes(x=c_age_years, y=uniqueness)) + 
    geom_point(fill="blue") + 
    ylim(0,1) +
  ylab("uniqueness") +
    theme_bw(base_size = 20) +
    theme(axis.title.x=element_blank(),
          axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
          plot.title = element_text(hjust = 0.5),
          legend.position = "none") +
    ggtitle("Phylum")


#Plot together in one panel
ggpubr::ggarrange(p.p_age, p.c_age, p.o_age, p.f_age, p.g_age, p.s_age, 
          ncol = 6, nrow = 1)

```


# PCOA of beta diversity distances across taxonomic levels

```{r}
#species
pcoa.species <- ape::pcoa(beta.div.species)
pcoa.species_variance_explained <- round(pcoa.species$values$Eigenvalues/sum(pcoa.species$values$Eigenvalues) * 100, 2)
pcoa.species.join <- pcoa.species$vectors[, 1:4] %>% as_tibble(rownames="sample.ID") %>%
  dplyr::rename(PCOA_1=Axis.1, PCOA_2=Axis.2, PCOA_3=Axis.3, PCOA_4=Axis.4) %>%
  dplyr::inner_join(sample.meta, by=c("sample.ID"="sample.ID")) 
#pcoa.species.join$dissimilar <- pcoa.species.join$sample.ID %in% sample.diss.species.list$sample.ID

#genus
pcoa.genus <- ape::pcoa(beta.div.genus)
pcoa.genus_variance_explained <- round(pcoa.genus$values$Eigenvalues/sum(pcoa.genus$values$Eigenvalues) * 100, 2)
pcoa.genus.join <- pcoa.genus$vectors[, 1:4] %>% as_tibble(rownames="sample.ID") %>%
  dplyr::rename(PCOA_1=Axis.1, PCOA_2=Axis.2, PCOA_3=Axis.3, PCOA_4=Axis.4) %>%
  dplyr::inner_join(sample.meta, by=c("sample.ID"="sample.ID"))
#pcoa.genus.join$dissimilar <- pcoa.genus.join$sample.ID %in% sample.diss.genus.list$sample.ID

#family
pcoa.family <- ape::pcoa(beta.div.family)
pcoa.family_variance_explained <- round(pcoa.family$values$Eigenvalues/sum(pcoa.family$values$Eigenvalues) * 100, 2)
pcoa.family.join <- pcoa.family$vectors[, 1:4] %>% as_tibble(rownames="sample.ID") %>%
  dplyr::rename(PCOA_1=Axis.1, PCOA_2=Axis.2, PCOA_3=Axis.3, PCOA_4=Axis.4) %>%
  dplyr::inner_join(sample.meta, by=c("sample.ID"="sample.ID"))
#pcoa.family.join$dissimilar <- pcoa.family.join$sample.ID %in% sample.diss.family.list$sample.ID

#order
pcoa.order <- ape::pcoa(beta.div.order)
pcoa.order_variance_explained <- round(pcoa.order$values$Eigenvalues/sum(pcoa.order$values$Eigenvalues) * 100, 2)
pcoa.order.join <- pcoa.order$vectors[, 1:4] %>% as_tibble(rownames="sample.ID") %>%
  dplyr::rename(PCOA_1=Axis.1, PCOA_2=Axis.2, PCOA_3=Axis.3, PCOA_4=Axis.4) %>%
  dplyr::inner_join(sample.meta, by=c("sample.ID"="sample.ID"))
#pcoa.order.join$dissimilar <- pcoa.order.join$sample.ID %in% sample.diss.order.list$sample.ID

#class
pcoa.class <- ape::pcoa(beta.div.class)
pcoa.class_variance_explained <- round(pcoa.class$values$Eigenvalues/sum(pcoa.class$values$Eigenvalues) * 100, 2)
pcoa.class.join <- pcoa.class$vectors[, 1:4] %>% as_tibble(rownames="sample.ID") %>%
  dplyr::rename(PCOA_1=Axis.1, PCOA_2=Axis.2, PCOA_3=Axis.3, PCOA_4=Axis.4) %>%
  dplyr::inner_join(sample.meta, by=c("sample.ID"="sample.ID"))
#pcoa.class.join$dissimilar <- pcoa.class.join$sample.ID %in% sample.diss.class.list$sample.ID

#phylum
pcoa.phylum <- ape::pcoa(beta.div.phylum)
pcoa.phylum_variance_explained <- round(pcoa.phylum$values$Eigenvalues/sum(pcoa.phylum$values$Eigenvalues) * 100, 2)
pcoa.phylum.join <- pcoa.phylum$vectors[, 1:4] %>% as_tibble(rownames="sample.ID") %>%
  dplyr::rename(PCOA_1=Axis.1, PCOA_2=Axis.2, PCOA_3=Axis.3, PCOA_4=Axis.4) %>%
  dplyr::inner_join(sample.meta, by=c("sample.ID"="sample.ID"))
#pcoa.phylum.join$dissimilar <- pcoa.phylum.join$sample.ID %in% sample.diss.phylum.list$sample.ID

```

## PCOA: PC1 and PC2

```{r, fig.dim=c(20,15)}
p.species <- ggplot(pcoa.species.join, aes(PCOA_1, PCOA_2))+
  geom_point(size=3, ) + 
  theme_bw(base_size=20) +
    theme(plot.title = element_text(hjust = 0.5),
          legend.position = "none") +
    ggtitle("Species") +
  xlab(paste0("PCOA_1 (", pcoa.species_variance_explained[1], "%)")) +
  ylab(paste0("PCOA_2 (", pcoa.species_variance_explained[2], "%)"))

    
p.genus <- ggplot(pcoa.genus.join, aes(PCOA_1, PCOA_2))+
  geom_point(size=3) + 
  theme_bw(base_size=20) +
    theme(plot.title = element_text(hjust = 0.5),
          legend.position = "none") +
    ggtitle("Genus") +
  xlab(paste0("PCOA_1 (", pcoa.genus_variance_explained[1], "%)")) +
  ylab(paste0("PCOA_2 (", pcoa.genus_variance_explained[2], "%)"))


p.family <- ggplot(pcoa.family.join, aes(PCOA_1, PCOA_2))+
  geom_point(size=3) + 
  theme_bw(base_size=20) +
    theme(plot.title = element_text(hjust = 0.5),
          legend.position = "none") +
    ggtitle("Family") +
  xlab(paste0("PCOA_1 (", pcoa.family_variance_explained[1], "%)")) +
  ylab(paste0("PCOA_2 (", pcoa.family_variance_explained[2], "%)"))


p.order <- ggplot(pcoa.order.join, aes(PCOA_1, PCOA_2))+
  geom_point(size=3) + 
  theme_bw(base_size=20) +
    theme(plot.title = element_text(hjust = 0.5),
          legend.position = "none") +
    ggtitle("Order") +
  xlab(paste0("PCOA_1 (", pcoa.order_variance_explained[1], "%)")) +
  ylab(paste0("PCOA_2 (", pcoa.order_variance_explained[2], "%)"))



p.class <- ggplot(pcoa.class.join, aes(PCOA_1, PCOA_2))+
  geom_point(size=3) + 
  theme_bw(base_size=20) +
    theme(plot.title = element_text(hjust = 0.5),
          legend.position = "none") +
    ggtitle("Class") +
  xlab(paste0("PCOA_1 (", pcoa.class_variance_explained[1], "%)")) +
  ylab(paste0("PCOA_2 (", pcoa.class_variance_explained[2], "%)"))



p.phylum <- ggplot(pcoa.phylum.join, aes(PCOA_1, PCOA_2))+
  geom_point(size=3) + 
  theme_bw(base_size=20) +
    theme(plot.title = element_text(hjust = 0.5),
          legend.position = "none") +
    ggtitle("Phylum") +
  xlab(paste0("PCOA_1 (", pcoa.phylum_variance_explained[1], "%)")) +
  ylab(paste0("PCOA_2 (", pcoa.phylum_variance_explained[2], "%)"))



ggpubr::ggarrange(p.phylum, p.class, p.order, p.family, p.genus, p.species, ncol = 3, nrow = 2)
```


## Save pcoa analysis results for PC1 and PC2

```{r}
#save pcoa information to excel file and rds
pcoa_list <- list(
  pcoa.phylum = pcoa.phylum.join %>% dplyr::select(sample.ID, c_age_years,PCOA_1, PCOA_2) %>%
  dplyr::mutate(age = case_when(c_age_years < 70 ~ "<70",
                                c_age_years >=70 & c_age_years <= 79 ~ "70-79",
                                c_age_years >=80 & c_age_years <= 99 ~ "80-99",
                                c_age_years >= 100 ~ ">100")) %>%
    dplyr::select(sample.ID, age, PCOA_1, PCOA_2),
  pcoa.class = pcoa.class.join %>% dplyr::select(sample.ID, PCOA_1, PCOA_2),
  pcoa.order = pcoa.order.join %>% dplyr::select(sample.ID, PCOA_1, PCOA_2),
  pcoa.family = pcoa.family.join %>% dplyr::select(sample.ID, PCOA_1, PCOA_2),
  pcoa.genus = pcoa.genus.join %>% dplyr::select(sample.ID, PCOA_1, PCOA_2),
  pcoa.species = pcoa.species.join %>% dplyr::select(sample.ID, PCOA_1, PCOA_2)
)
saveRDS(pcoa_list, paste0(outpath,"pcoa_bracken_chinese.R1.K3.rds"))
write_xlsx(pcoa_list, paste0(outpath,"pcoa_bracken_chinese.R1.K3.xlsx"))
```


## PERMANOVA: Association with age

```{r}
samp.dat <- as.matrix(subset_chinese_biom@sam_data)
#species level
dist_matrix.species <- as.matrix(beta.div.species)
samp.dat.species <- as.data.frame(samp.dat[rownames(dist_matrix.species),])

pm.species<-adonis2(dist_matrix.species~as.integer(samp.dat.species$c_age_years), data=samp.dat.species, permutations=999, by="margin")
pm.species

pval.species <- pm.species$`Pr(>F)`[1]
F.species <- pm.species$F[1] %>% signif(3)

#genus level
dist_matrix.genus <- as.matrix(beta.div.genus)
samp.dat.genus <- as.data.frame(samp.dat[rownames(dist_matrix.genus),])

pm.genus<-adonis2(dist_matrix.genus~as.integer(samp.dat.genus$c_age_years), data=samp.dat.genus, permutations=999, by="margin")
pm.genus

pval.genus <- pm.genus$`Pr(>F)`[1]
F.genus <- pm.genus$F[1] %>% signif(3)

#family level
dist_matrix.family <- as.matrix(beta.div.family)
samp.dat.family <- as.data.frame(samp.dat[rownames(dist_matrix.family),])

pm.family<-adonis2(dist_matrix.family~as.integer(samp.dat.family$c_age_years), data=samp.dat.family, permutations=999, by="margin")
pm.family

pval.family <- pm.family$`Pr(>F)`[1]
F.family <- pm.family$F[1] %>% signif(3)

#order level
dist_matrix.order <- as.matrix(beta.div.order)
samp.dat.order <- as.data.frame(samp.dat[rownames(dist_matrix.order),])

pm.order<-adonis2(dist_matrix.order~as.integer(samp.dat.order$c_age_years), data=samp.dat.order, permutations=999, by="margin")
pm.order

pval.order <- pm.order$`Pr(>F)`[1]
F.order <- pm.order$F[1] %>% signif(3)

#class level
dist_matrix.class <- as.matrix(beta.div.class)
samp.dat.class <- as.data.frame(samp.dat[rownames(dist_matrix.class),])

pm.class<-adonis2(dist_matrix.class~as.integer(samp.dat.class$c_age_years), data=samp.dat.class, permutations=999, by="margin")
pm.class

pval.class <- pm.class$`Pr(>F)`[1]
F.class <- pm.class$F[1] %>% signif(3)

#phylum level
dist_matrix.phylum <- as.matrix(beta.div.phylum)
samp.dat.phylum <- as.data.frame(samp.dat[rownames(dist_matrix.phylum),])

pm.phylum<-adonis2(dist_matrix.phylum~as.integer(samp.dat.phylum$c_age_years), data=samp.dat.phylum, permutations=999, by="margin")
pm.phylum

pval.phylum <- pm.phylum$`Pr(>F)`[1]
F.phylum <- pm.phylum$F[1] %>% signif(3)

```


## PCOA colored by age (continuous)

```{r, fig.dim=c(30,20)}
p.plot <- ggplot(pcoa.species.join, aes(PCOA_1, PCOA_2))+
  geom_point(aes(colour=c_age_years), size=6) + 
  theme_bw(base_size=60) +
    theme(plot.title = element_text(hjust = 0.5), 
          legend.key.height= unit(1.8, 'cm'),
        legend.key.width= unit(1, 'cm')) +
  scale_color_continuous(name = "age", type="viridis", limits=c(50,110)) 

p.legend <- p.plot %>% get_legend() %>% as_ggplot

p.species <- ggplot(pcoa.species.join, aes(PCOA_1, PCOA_2, color=age_group))+
  geom_point(aes(colour=c_age_years), size=6) + 
  theme_bw(base_size=50) +
    theme(plot.title = element_text(hjust = 0.5),
          legend.position = "none") +
    labs(title="Species",
         subtitle=paste(paste0("F=",F.species,","),paste0("p=",pval.species))) +
  scale_color_continuous(type="viridis", limits=c(50,110)) +
  xlab(paste0("PCOA_1 (", pcoa.species_variance_explained[1], "%)")) +
  ylab(paste0("PCOA_2 (", pcoa.species_variance_explained[2], "%)")) +
  ylim(-0.6,0.6) +
  xlim(-0.6,0.5)
    
p.genus <- ggplot(pcoa.genus.join, aes(PCOA_1, PCOA_2, color=age_group))+
  geom_point(aes(colour=c_age_years), size=6) + 
  theme_bw(base_size=50) +
    theme(plot.title = element_text(hjust = 0.5),
          legend.position = "none") +
    labs(title="Genus",
         subtitle=paste(paste0("F=",F.genus,","),paste0("p=",pval.genus))) +
  scale_color_continuous(type="viridis", limits=c(50,110)) +
  xlab(paste0("PCOA_1 (", pcoa.genus_variance_explained[1], "%)")) +
  ylab(paste0("PCOA_2 (", pcoa.genus_variance_explained[2], "%)")) +
  ylim(-0.6,0.6) +
  xlim(-0.6,0.6)

p.family <- ggplot(pcoa.family.join, aes(PCOA_1, PCOA_2))+
  geom_point(aes(colour=c_age_years), size=6) + 
  theme_bw(base_size=50) +
    theme(plot.title = element_text(hjust = 0.5),
          legend.position = "none") +
    labs(title="Family",
         subtitle=paste(paste0("F=",F.family,","),paste0("p=",pval.family))) +
  scale_color_continuous(type="viridis", limits=c(50,110)) +
  xlab(paste0("PCOA_1 (", pcoa.family_variance_explained[1], "%)")) +
  ylab(paste0("PCOA_2 (", pcoa.family_variance_explained[2], "%)")) +
  ylim(-0.6,0.8) +
  xlim(-0.6,0.6)

p.order <- ggplot(pcoa.order.join, aes(PCOA_1, PCOA_2))+
  geom_point(aes(colour=c_age_years), size=6) + 
  theme_bw(base_size=50) +
    theme(plot.title = element_text(hjust = 0.5),
          legend.position = "none") +
    labs(title="Order",
         subtitle=paste(paste0("F=",F.order,","),paste0("p=",pval.order))) +
  scale_color_continuous(type="viridis", limits=c(50,110)) +
  xlab(paste0("PCOA_1 (", pcoa.order_variance_explained[1], "%)")) +
  ylab(paste0("PCOA_2 (", pcoa.order_variance_explained[2], "%)")) +
 ylim(-0.8,1) +
  xlim(-0.7,0.8)

p.class <- ggplot(pcoa.class.join, aes(PCOA_1, PCOA_2))+
  geom_point(aes(colour=c_age_years), size=6) + 
  theme_bw(base_size=50) +
    theme(plot.title = element_text(hjust = 0.5),
          legend.position = "none") +
    labs(title="Class",
         subtitle=paste(paste0("F=",F.class,","),paste0("p=",pval.class))) +
  scale_color_continuous(type="viridis", limits=c(50,110)) +
  xlab(paste0("PCOA_1 (", pcoa.class_variance_explained[1], "%)")) +
  ylab(paste0("PCOA_2 (", pcoa.class_variance_explained[2], "%)")) +
  ylim(-0.8,1) +
  xlim(-0.7,0.8)

p.phylum <- ggplot(pcoa.phylum.join, aes(PCOA_1, PCOA_2))+
  geom_point(aes(colour=c_age_years), size=6) + 
  theme_bw(base_size=50) +
    theme(plot.title = element_text(hjust = 0.5),
          legend.position = "none") +
    labs(title="Phylum",
         subtitle=paste(paste0("F=",F.phylum,","),paste0("p=",pval.phylum))) +
 scale_color_continuous(type="viridis", limits=c(50,110)) +
  xlab(paste0("PCOA_1 (", pcoa.phylum_variance_explained[1], "%)")) +
  ylab(paste0("PCOA_2 (", pcoa.phylum_variance_explained[2], "%)")) +
  ylim(-0.8,0.6) +
  xlim(-0.5,0.8)


#Plot together in one panel
ggpubr::ggarrange(p.phylum, p.class, p.order, p.legend, p.family, p.genus, p.species, ncol = 4, nrow = 2)

#create directory for figures
dir.create(paste0(outpath, "Figures/bracken_R1_K3/"))

#save figures
png(paste0(outpath,"Figures/bracken_R1_K3/pcoa_bracken_R1_K3_chinese_alllevels.png"), units="in", width=50, height=20, res=300)
ggpubr::ggarrange(p.phylum, p.class, p.order, p.legend, p.family, p.genus, p.species, ncol = 4, nrow = 2)
dev.off()

png(paste0(outpath,"Figures/bracken_R1_K3/pcoa_bracken_R1_K3_chinese_main.png"), units="in", width=40, height=12, res=300)
ggpubr::ggarrange(p.phylum, p.genus, p.species, p.legend, ncol = 4, nrow = 1)
dev.off()

png(paste0(outpath,"Figures/bracken_R1_K3/pcoa_bracken_R1_K3_chinese_supp.png"), units="in", width=40, height=12, res=300)
ggpubr::ggarrange(p.class, p.order, p.family, p.legend, ncol = 4, nrow = 1)
dev.off()

```

```{r, fig.dim=c(30,20)}
#save each figure separately
p.species <- ggplot(pcoa.species.join, aes(PCOA_1, PCOA_2, color=age_group))+
  geom_point(aes(colour=c_age_years), size=3) + 
  theme_bw(base_size=35) +
    theme(plot.title = element_text(hjust = 0.5), 
          legend.key.height= unit(1, 'cm'),
        legend.key.width= unit(1, 'cm')) +
    labs(title="Species" , subtitle=paste0("p=",pval.species)) +
  scale_color_continuous(name = "age", type="viridis", limits=c(50,110)) +
  xlab(paste0("PCOA_1 (", pcoa.species_variance_explained[1], "%)")) +
  ylab(paste0("PCOA_2 (", pcoa.species_variance_explained[2], "%)")) +
   ylim(-0.6,0.6) +
  xlim(-0.6,0.5)
    
p.genus <- ggplot(pcoa.genus.join, aes(PCOA_1, PCOA_2, color=age_group))+
  geom_point(aes(colour=c_age_years), size=3) + 
  theme_bw(base_size=35) +
    theme(plot.title = element_text(hjust = 0.5), 
          legend.key.height= unit(1, 'cm'),
        legend.key.width= unit(1, 'cm')) +
    labs(title="Genus", subtitle=paste0("p=",pval.genus)) +
  scale_color_continuous(name = "age", type="viridis", limits=c(50,110)) +
  xlab(paste0("PCOA_1 (", pcoa.genus_variance_explained[1], "%)")) +
  ylab(paste0("PCOA_2 (", pcoa.genus_variance_explained[2], "%)")) +
  ylim(-0.6,0.6) +
  xlim(-0.6,0.6)

p.family <- ggplot(pcoa.family.join, aes(PCOA_1, PCOA_2))+
  geom_point(aes(colour=c_age_years), size=3) + 
  theme_bw(base_size=35) +
    theme(plot.title = element_text(hjust = 0.5), 
          legend.key.height= unit(1, 'cm'),
        legend.key.width= unit(1, 'cm')) +
    labs(title="Family", subtitle=paste0("p=",pval.family)) +
  scale_color_continuous(name = "age", type="viridis", limits=c(50,110)) +
  xlab(paste0("PCOA_1 (", pcoa.family_variance_explained[1], "%)")) +
  ylab(paste0("PCOA_2 (", pcoa.family_variance_explained[2], "%)")) +
  ylim(-0.6,0.8) +
  xlim(-0.6,0.6)

p.order <- ggplot(pcoa.order.join, aes(PCOA_1, PCOA_2))+
  geom_point(aes(colour=c_age_years), size=3) + 
  theme_bw(base_size=35) +
    theme(plot.title = element_text(hjust = 0.5), 
          legend.key.height= unit(1, 'cm'),
        legend.key.width= unit(1, 'cm')) +
    labs(title="Order", subtitle=paste0("p=",pval.order)) +
  scale_color_continuous(name = "age", type="viridis", limits=c(50,110)) +
  xlab(paste0("PCOA_1 (", pcoa.order_variance_explained[1], "%)")) +
  ylab(paste0("PCOA_2 (", pcoa.order_variance_explained[2], "%)")) +
  ylim(-0.8,1) +
  xlim(-0.7,0.8)

p.class <- ggplot(pcoa.class.join, aes(PCOA_1, PCOA_2))+
  geom_point(aes(colour=c_age_years), size=3) + 
  theme_bw(base_size=35) +
    theme(plot.title = element_text(hjust = 0.5), 
          legend.key.height= unit(1, 'cm'),
        legend.key.width= unit(1, 'cm')) +
    labs(title="Class", subtitle=paste0("p=",pval.class)) +
  scale_color_continuous(name = "age", type="viridis", limits=c(50,110)) +
  xlab(paste0("PCOA_1 (", pcoa.class_variance_explained[1], "%)")) +
  ylab(paste0("PCOA_2 (", pcoa.class_variance_explained[2], "%)")) +
  ylim(-0.8,1) +
  xlim(-0.7,0.8)

p.phylum <- ggplot(pcoa.phylum.join, aes(PCOA_1, PCOA_2))+
  geom_point(aes(colour=c_age_years), size=3) + 
  theme_bw(base_size=35) +
    theme(plot.title = element_text(hjust = 0.5), 
          legend.key.height= unit(1, 'cm'),
        legend.key.width= unit(1, 'cm')) +
    labs(title="Phylum", subtitle=paste0("p=",pval.phylum)) +
 scale_color_continuous(name = "age", type="viridis", limits=c(50,110)) +
  xlab(paste0("PCOA_1 (", pcoa.phylum_variance_explained[1], "%)")) +
  ylab(paste0("PCOA_2 (", pcoa.phylum_variance_explained[2], "%)")) +
  ylim(-0.8,0.5) +
  xlim(-0.5,0.8)

#save each plot
png(paste0(outpath,"Figures/bracken_R1_K3/pcoa_bracken_R1_K3_chinese_phylum.png"), units="in", width=12, height=10, res=300)
p.phylum
dev.off()

png(paste0(outpath,"Figures/bracken_R1_K3/pcoa_bracken_R1_K3_chinese_class.png"), units="in", width=12, height=10, res=300)
p.class
dev.off()

png(paste0(outpath,"Figures/bracken_R1_K3/pcoa_bracken_R1_K3_chinese_order.png"), units="in", width=12, height=10, res=300)
p.order
dev.off()

png(paste0(outpath,"Figures/bracken_R1_K3/pcoa_bracken_R1_K3_chinese_family.png"), units="in", width=12, height=10, res=300)
p.family
dev.off()

png(paste0(outpath,"Figures/bracken_R1_K3/pcoa_bracken_R1_K3_chinese_genus.png"), units="in", width=12, height=10, res=300)
p.genus
dev.off()

png(paste0(outpath,"Figures/bracken_R1_K3/pcoa_bracken_R1_K3_chinese_species.png"), units="in", width=12, height=10, res=300)
p.species
dev.off()
```
