---
title: "Alpha diversity analysis with Xu et al chinese cohort data based in-common taxa between taxonomic classification methods"
author: "Tanya Karagiannis"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    theme: paper
    toc: yes
    toc_collapse: no
    toc_float: yes
  pdf_document:
    toc: yes
---

```{r setup, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
library(dplyr)
library(ggplot2)
library(ggpubr)
library(ape)
library(DESeq2)
library(stringr)
library(reshape2)
library(graphics)

library(codebook)
library(future)
#library(vtable)
library(useful)

library(vegan)
library(data.table)

# Data Cleaning
library(janitor)

# EDA
library(skimr)
library(DataExplorer)

# ggplot2 Helpers
library(ggplot2)
library(ggsignif)
library(scales)
library(phyloseq)
library(tidyverse)
library(tidymodels)
#heatmap
library(pheatmap)

#specify file path
flpath <- "/restricted/projectnb/uh2-sebas/data/metagenomics/external_cohorts/Han_Chinese_Centenarian/processed_data/data_library/"
outpath <- "/restricted/projectnb/uh2-sebas/analysis/metagenomics/parallel_systems_comparison/03_alpha_diversity_analysis/Chinese_cohort/output/"
meta.dir <- "/restricted/projectnb/necs/Data_library/Metabolomics/resources/Xu_plasma_metabolite_analysis/"
sra.dir <- "/restricted/projectnb/uh2-sebas/data/metagenomics/external_cohorts/Han_Chinese_Centenarian/metadata/"

#metaphlan data
metaphlan_otufl <- "metaphlan4_renorm_chinese_phyloseq.10.23.2024.rds"
#bracken data with log10 read count threshold 1 and unique kmer minimizer count threshold 3
bracken_otufl <- "bracken_renorm_chinese_phyloseq.R1.K3.10.22.2024.rds"

#metadata
phenofl <- "phenotype.xlsx"

#set seed
set.seed(157)

```

This script is for diversity analysis using data after QC.

```{r user-defined functions}
## outlier function
is_outlier <- function(x) {
  return(x < quantile(x, 0.25) - 1.5 * IQR(x) | x > quantile(x, 0.75) + 1.5 * IQR(x))
}


# calculate relative abundance/normalized abundance
get_relabund <- function(indat, varname){
  relabund <- as.tibble(indat) %>%
    janitor::adorn_percentages("col") %>%
    column_to_rownames(var = varname) 
}
# 
source("/restricted/projectnb/uh2-sebas/analysis/metagenomics/ye_analyses/Scripts/UserDefinedFunctions/CTDS_function.R")

#function to aggregate otu abundances, calculate proportions and calculated CTDS based on taxonomic hierarchy
otu_ctds <- function(taxadata = biom_chinese_kraken.taxa, 
                     otudata=biom_chinese_kraken.otu, rank=rank, 
                     sample.meta=sample.meta){
  #load ctds function
  source("/restricted/projectnb/uh2-sebas/analysis/metagenomics/ye_analyses/Scripts/UserDefinedFunctions/CTDS_function.R")
  
  #prepare taxonomic and otu data to run diversity analysis
  #convert taxonomic table to a tibble with rownames as a column called tax.ID
  #filter taxonomic table to remove taxa with no species label
  biom_taxdata <- as_tibble(taxadata, rownames = "tax.ID") 
 # message(paste("taxa table rows:",nrow(biom_taxdata)))
  #convert otu table to a tibble with rownames as a column called tax.ID
  biom_otudata <- as_tibble(otudata, rownames = "tax.ID")
  
  #save sample names
  biom_samplenames <- colnames(otudata)
  #message(paste("taxa table rows:",nrow(biom_otudata)))

  
  #taxa dictionary: taxonomy table with taxa ID and taxa names
  taxa_dict <- biom_taxdata %>% 
    dplyr::select(tax.ID, {{rank}}) %>%
    unite("tax.name", {{rank}}, remove = FALSE)

  #use an inner_join function from dplyr to join data from both table based on common tax ID
  #create new variable taxanames to create unique
  biom_join <- dplyr::inner_join(taxa_dict, biom_otudata, by = "tax.ID")
  #message(paste("joined table rows:", nrow(biom_join))) 


  #remove rankings and transpose data for samples in rows and taxa IDs in columns
  #make sure the samples are set as rownames
  # biom_prop <- biom_join %>% 
  #   group_by(tax.name) %>% #use rank groupings
  #   summarise_if(is.numeric, sum) %>% #aggregate OTU abundances for each taxa in the rank
  #   select(c(tax.name, contains(".aggregrated.report"))) %>% #select rank and sample OTU abundances
  #   dplyr::rename(tax.ID = tax.name) %>%
  #   tibble::column_to_rownames(var = "tax.ID") %>% #make tax.ID variable to rownames
  #   t()
  
  biom_prop <- biom_join %>% 
    group_by(tax.name) %>% #use rank groupings
    summarise_if(is.numeric, sum) %>% #aggregate OTU abundances for each taxa in the rank
    dplyr::select(tax.name, biom_samplenames) %>% #select rank and sample OTU abundances
    column_to_rownames(var = "tax.name") %>% #make tax.name variable to rownames
    t() %>% #transpose data to have taxa in columns and samples as rows
    as_tibble(rownames = "sample.ID") %>% #make transposed data into a tibble with rownames as sample.ID variable
    #janitor::adorn_percentages() %>% #use janitor package to get row proportions for each sample
    column_to_rownames(var = "sample.ID") #add sample IDs back to rownames
  

  #ctds function
  biom.res <- CTDS.score(biom_prop, metadata = sample.meta)
  return(biom.res)
  

}


prop_mat <- function(taxadata = biom_chinese_kraken.taxa, 
                     otudata=biom_chinese_kraken.otu, rank=rank, 
                     sample.meta=sample.meta){

  #convert taxonomic table to a tibble with rownames as a column called tax.ID
  #filter taxonomic table to remove taxa with no taxa classification in any level
  biom_taxdata <- as_tibble(taxadata, rownames = "tax.ID") %>% 
    unite("tax.name", kingdom:{{rank}}, remove = FALSE)

  #row number for taxa data after removing incomplete classification
  message(paste("taxa table rows:",nrow(biom_taxdata)))

  #convert otu table to a tibble with rownames as a column called tax.ID
  biom_otudata <- as_tibble(otudata, rownames = "tax.ID")
  #row number for otu data before removing incomplete classification
  message(paste("otu table rows:",nrow(biom_otudata)))

  #use an inner_join function from dplyr to join data from both table based on common tax ID
  biom_join <- inner_join(biom_taxdata, biom_otudata, by = "tax.ID")
  #row number for joined data
  message(paste("joined table rows:", nrow(biom_join))) #row number 11039 for the common taxa
  
  biom_taxid <- biom_join %>% select(c(tax.ID, tax.name))
  

  #remove rankings and transpose data for samples in rows and taxa IDs in columns
  #make sure the samples are set as rownames
  biom_prop <- biom_join %>% 
    group_by(tax.name) %>%
    summarise_if(is.numeric, sum) %>% 
    select(c(tax.name, contains(".aggregrated.report"))) %>%
    column_to_rownames(var = "tax.name") %>% #make tax.name variable to rownames
    t() %>% #transpose data to have taxa in columns and samples as rows
    as_tibble(rownames = "sample.ID") %>% #make transposed data into a tibble with rownames as sample.ID variable
    janitor::adorn_percentages() %>% #use janitor package to get row proportions for each sample
    column_to_rownames(var = "sample.ID") #add sample IDs back to rownames
  dim(biom_prop)
  useful::corner(biom_prop) #use this to visualize a corner of the data

  #rerun abundance table
  biom.res <- CTDS.score(biom_prop, metadata = sample.meta)
  return(biom.res)

}
```

# Read in phenotypic data


```{r, message=FALSE}
# Extract phenotypic data
#load clinical metadata information
#clinical.data <- read_excel(paste0(meta.dir, "DifferentialAbundance_Supplement.xlsx"), sheet = "Table S1 Metadata", skip = 1)
clinical.data <- read_excel(paste0(meta.dir, phenofl))

#add ethnicity variable
clinical.data$ethnicity <- "han_chinese"


#extract sample IDs and create family ID and group ID
sample.ID <- clinical.data %>% dplyr::select(1) %>% pull()

fam.group <- str_remove(sample.ID, "QCS-") %>%                       
              str_split("-") %>%
                unlist() 

fam.group <- matrix(fam.group, nrow=2)

#create family ID
fam.id <- fam.group[1,]

#create group ID
group.id <- fam.group[2,]
group.id[group.id == "1"] <- "Advanced Age"
group.id[group.id == "2" | group.id == "3"] <- "Offspring"


#add to clinical data
clinical.clean <- clinical.data %>%
  mutate(sample.ID = sample.ID, fam.id = fam.id, group = group.id) 
clinical.clean$"C-age (years)" <- as.integer(clinical.clean$"C-age (years)")


#load sample Run names and sample SRA names
SRA.data <- read_csv(paste0(sra.dir,"SraRunInfo.csv"))

SRA.clean <- SRA.data %>% 
  dplyr::select(Run, SampleName) 

#sample ID and sample SRA names
sample.dict <- read_excel(paste0(sra.dir, "sample name-sequencing ID sheet.xlsx"))

#combine files
sample.info <- dplyr::inner_join(SRA.clean, sample.dict, by = c("SampleName"))

pheno.data <- dplyr::inner_join(sample.info, clinical.clean, by = c("sample.ID")) %>% 
                                  dplyr::select(-c(sample.ID)) %>%
                                  mutate(stool_kit = Run, Sample=Run) %>%
                                  dplyr::rename(c_age_years="C-age (years)")
pheno.data$stool_kit <- as.character(pheno.data$stool_kit)
pheno.data$Sample <- as.character(pheno.data$Sample)


SAMPLE_phy <- sample_data(pheno.data %>% column_to_rownames(var="Sample"))
```


# Metaphlan4: Read in phenotype data and phyloseq object
```{r metaphlan data}
#load phyloseq obj
metaphlan_chinese_biom <- readRDS(paste0(flpath, metaphlan_otufl))

#add phenotypic data to phyloseq object
metaphlan_chinese_biom@sam_data <- SAMPLE_phy

#keep samples with age group label
metaphlan_chinese_biom <- subset_samples(metaphlan_chinese_biom, !is.na(c_age_years))

metaphlan.sample.meta <- sample_data(metaphlan_chinese_biom) %>% 
  as_tibble(rownames="sample.ID")
# metaphlan.samples <- metaphlan.sample.meta %>% dplyr::select(sample.ID, stool_kit)
  
OTU_metaphlan <- metaphlan_chinese_biom@otu_table@.Data
TAX_metaphlan <- metaphlan_chinese_biom@tax_table@.Data

#save sample data for permanova
# rownames(metaphlan_chinese_biom@sam_data) <- metaphlan_chinese_biom@sam_data$stool_kit
# sample_names(metaphlan_chinese_biom) <- metaphlan_chinese_biom@sam_data$stool_kit
metaphlan.samp.dat <- data.frame(metaphlan_chinese_biom@sam_data)

```

# Bracken: Read in phenotype data and phyloseq object
```{r bracken data}
#load phyloseq obj
bracken_chinese_biom <- readRDS(paste0(flpath, bracken_otufl))

#add phenotypic data to phyloseq object
bracken_chinese_biom@sam_data <- SAMPLE_phy

#keep samples with age group label
bracken_chinese_biom <- subset_samples(bracken_chinese_biom, !is.na(c_age_years))

bracken.sample.meta <- sample_data(bracken_chinese_biom) %>%
  as_tibble(rownames="sample.ID")
# bracken.samples <- bracken.sample.meta %>% dplyr::select(sample.ID, stool_kit)

#calculate relative abundances
OTU_bracken <- bracken_chinese_biom@otu_table@.Data
TAX_bracken <- bracken_chinese_biom@tax_table@.Data

#save sample data for permanova
# rownames(bracken_chinese_biom@sam_data) <- bracken_chinese_biom@sam_data$stool_kit
# sample_names(bracken_chinese_biom) <- bracken_chinese_biom@sam_data$stool_kit
bracken.samp.dat <- data.frame(bracken_chinese_biom@sam_data)

```

# common samples

```{r}
#metaphlan
metaphlan.sample.meta <- metaphlan.sample.meta %>% dplyr::filter(metaphlan.sample.meta$sample.ID %in% bracken.sample.meta$sample.ID)
dim(metaphlan.sample.meta)

OTU_metaphlan <- OTU_metaphlan[, metaphlan.sample.meta$sample.ID]
dim(OTU_metaphlan)

#bracken
bracken.sample.meta <- bracken.sample.meta %>% dplyr::filter(bracken.sample.meta$sample.ID %in% metaphlan.sample.meta$sample.ID)
dim(bracken.sample.meta)

OTU_bracken <- OTU_bracken[, bracken.sample.meta$sample.ID]
dim(OTU_bracken)
```




# Filter to common taxa and renormalize data

```{r common taxa}
#adjust taxa table
#move rownames to column variable tax.ID
metaphlan_chinese_biom.taxa <- metaphlan_chinese_biom@tax_table@.Data %>% as_tibble(rownames = "tax.ID")
bracken_chinese_biom.taxa <- bracken_chinese_biom@tax_table@.Data %>% as_tibble(rownames = "tax.ID")

#combine taxa tables based keeping common taxa using inner_join
common_taxa_total <- dplyr::inner_join(metaphlan_chinese_biom.taxa, bracken_chinese_biom.taxa, by= c("tax.ID")) 
dim(common_taxa_total)

metaphlan_taxa_total <- common_taxa_total %>% dplyr::select(tax.ID:Species) %>% column_to_rownames(var = "tax.ID")
bracken_taxa_total <- common_taxa_total %>% dplyr::select(tax.ID, kingdom:species) %>% column_to_rownames(var = "tax.ID")
dim(metaphlan_taxa_total)
dim(bracken_taxa_total)

#combine taxa tables based keeping common taxa using inner_join
filtered_taxa_total <- dplyr::inner_join(metaphlan_chinese_biom.taxa, bracken_chinese_biom.taxa, by= c("tax.ID")) %>%
  dplyr::select(tax.ID, kingdom:species) %>%
                  column_to_rownames(var = "tax.ID")
dim(filtered_taxa_total)

#subset otu table for common taxa
metaphlan_chinese_biom.otu <- OTU_metaphlan[rownames(metaphlan_taxa_total),] %>% as.matrix()
bracken_chinese_biom.otu <- OTU_bracken[rownames(bracken_taxa_total),] %>% as.matrix()

#plot relative abundances of common taxa before renormalizing
common.relabund.metaphlan <- colSums(metaphlan_chinese_biom.otu)
common.relabund.bracken <- colSums(bracken_chinese_biom.otu)

#renormalize metaphlan data
metaphlan_chinese_renorm.otu <- t(t(metaphlan_chinese_biom.otu)/colSums(metaphlan_chinese_biom.otu))
#renormalize bracken data
bracken_chinese_renorm.otu <- t(t(bracken_chinese_biom.otu)/colSums(bracken_chinese_biom.otu))
# colnames(metaphlan_chinese_renorm.otu) <- paste0(colnames(metaphlan_chinese_renorm.otu), "_Metaphlan")
# colnames(bracken_chinese_renorm.otu) <- paste0(colnames(bracken_chinese_renorm.otu), "_Bracken")

```


# Alpha diversity analysis using Adjusted Entropy 

## Metaphlan: Calculate alpha diversity using otu_ctds function

```{r, warning=FALSE}
#taxa diversity for species
OTU_species <- metaphlan_chinese_renorm.otu
metaphlan.ctds.species <- otu_ctds(taxadata = filtered_taxa_total, otudata = OTU_species, rank = c("kingdom","phylum","class","order","family","genus","species"), sample.meta = metaphlan.sample.meta)

#taxa diversity for genus
OTU_genus <- metaphlan_chinese_renorm.otu
metaphlan.ctds.genus <- otu_ctds(taxadata = filtered_taxa_total, otudata = OTU_genus, rank = c("kingdom","phylum","class","order","family","genus"), sample.meta = metaphlan.sample.meta)

#taxa diversity for family
OTU_family <- metaphlan_chinese_renorm.otu
metaphlan.ctds.family <- otu_ctds(taxadata = filtered_taxa_total, otudata = OTU_family, rank = c("kingdom","phylum","class","order","family"), sample.meta = metaphlan.sample.meta)

#taxa diversity for order
OTU_order <- metaphlan_chinese_renorm.otu
metaphlan.ctds.order <- otu_ctds(taxadata = filtered_taxa_total, otudata = OTU_order, rank = c("kingdom","phylum","class","order"), sample.meta = metaphlan.sample.meta)

#taxa diversity for class
OTU_class <- metaphlan_chinese_renorm.otu
metaphlan.ctds.class <- otu_ctds(taxadata = filtered_taxa_total, otudata = OTU_class, rank = c("kingdom","phylum","class"), sample.meta = metaphlan.sample.meta)

#taxa diversity for phylum
OTU_phylum <- metaphlan_chinese_renorm.otu
metaphlan.ctds.phylum <- otu_ctds(taxadata = filtered_taxa_total, otudata = OTU_phylum, rank = c("kingdom","phylum"), sample.meta = metaphlan.sample.meta)

#extract scores and save as csv file
metaphlan.ctds.phylum.score <- metaphlan.ctds.phylum %>% dplyr::select(sample.ID, statistic) %>% dplyr::rename(phylum_statistic=statistic)
metaphlan.ctds.class.score <- metaphlan.ctds.class %>% dplyr::select(statistic) %>% dplyr::rename(class_statistic=statistic)
metaphlan.ctds.order.score <- metaphlan.ctds.order %>% dplyr::select(statistic) %>% dplyr::rename(order_statistic=statistic)
metaphlan.ctds.family.score <- metaphlan.ctds.family %>% dplyr::select(statistic) %>% dplyr::rename(family_statistic=statistic)
metaphlan.ctds.genus.score <- metaphlan.ctds.genus %>% dplyr::select(statistic) %>% dplyr::rename(genus_statistic=statistic)
metaphlan.ctds.species.score <- metaphlan.ctds.species %>% dplyr::select(statistic) %>% dplyr::rename(species_statistic=statistic)

metaphlan.ctds.list <- dplyr::bind_cols(metaphlan.ctds.phylum.score, metaphlan.ctds.class.score, metaphlan.ctds.order.score, metaphlan.ctds.family.score, metaphlan.ctds.genus.score, metaphlan.ctds.species.score)

write_csv(metaphlan.ctds.list, file=paste0(outpath,"entropy.metaphlan.commontaxa.chinese.csv"))

```

### Age association

```{r}
#species level
lm.species <- lm(statistic ~ c_age_years, data=metaphlan.ctds.species)
pval.species <- broom::tidy(lm.species) %>% dplyr::filter(term=="c_age_years") %>% pull(p.value) %>% signif(3)
est.species <- broom::tidy(lm.species) %>% dplyr::filter(term=="c_age_years") %>% pull(estimate) %>% signif(3)
summary(lm.species)

#genus level
lm.genus <- lm(statistic ~ c_age_years, data=metaphlan.ctds.genus)
pval.genus <- broom::tidy(lm.genus) %>% dplyr::filter(term=="c_age_years") %>% pull(p.value) %>% signif(3)
est.genus <- broom::tidy(lm.genus) %>% dplyr::filter(term=="c_age_years") %>% pull(estimate) %>% signif(3)
summary(lm.genus)

#family level
lm.family <- lm(statistic ~ c_age_years, data=metaphlan.ctds.family)
pval.family <- broom::tidy(lm.family) %>% dplyr::filter(term=="c_age_years") %>% pull(p.value) %>% signif(3)
est.family <- broom::tidy(lm.family) %>% dplyr::filter(term=="c_age_years") %>% pull(estimate) %>% signif(3)
summary(lm.family)

#order level
lm.order <- lm(statistic ~ c_age_years, data=metaphlan.ctds.order)
pval.order <- broom::tidy(lm.order) %>% dplyr::filter(term=="c_age_years") %>% pull(p.value) %>% signif(3)
est.order <- broom::tidy(lm.order) %>% dplyr::filter(term=="c_age_years") %>% pull(estimate) %>% signif(3)
summary(lm.order)

#class level
lm.class <- lm(statistic ~ c_age_years, data=metaphlan.ctds.class)
pval.class <- broom::tidy(lm.class) %>% dplyr::filter(term=="c_age_years") %>% pull(p.value) %>% signif(3)
est.class <- broom::tidy(lm.class) %>% dplyr::filter(term=="c_age_years") %>% pull(estimate) %>% signif(3)
summary(lm.class)

#phylum level
lm.phylum <- lm(statistic ~ c_age_years, data=metaphlan.ctds.phylum)
pval.phylum <- broom::tidy(lm.phylum) %>% dplyr::filter(term=="c_age_years") %>% pull(p.value) %>% signif(3)
est.phylum <- broom::tidy(lm.phylum) %>% dplyr::filter(term=="c_age_years") %>% pull(estimate) %>% signif(3)
summary(lm.phylum)
```

### Plot Alpha diversity across taxonomic levels

```{r, fig.dim = c(50,10)}
#For species level
p.s_age <- ggplot(metaphlan.ctds.species, aes(x=c_age_years, y=statistic)) + 
    geom_point() + 
  geom_smooth(method="lm") +
    ylim(-1,0) + 
   xlim(50, 110) +
    theme_bw(base_size=40) +
    theme(axis.title.y=element_blank(),
          axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
          plot.title = element_text(hjust = 0.5),
          legend.position = "none") +
    ggtitle("Species", subtitle=paste(
      paste0("b=",est.species), 
      paste0("p=",pval.species)
      )) +
  xlab("age")
    

#For genus level
p.g_age <- ggplot(metaphlan.ctds.genus, aes(x=c_age_years, y=statistic)) + 
    geom_point() + 
  geom_smooth(method="lm") +
    ylim(-1,0) +
   xlim(50, 110) +
    theme_bw(base_size=40) +
    theme(axis.title.y=element_blank(),
          axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
          plot.title = element_text(hjust = 0.5),
          legend.position = "none") +
    ggtitle("Genus", subtitle=paste(
      paste0("b=",est.genus), 
      paste0("p=",pval.genus)
      )) +
  xlab("age")
    

#For family level
p.f_age <- ggplot(metaphlan.ctds.family, aes(x=c_age_years, y=statistic)) + 
    geom_point() + 
  geom_smooth(method="lm") +
  ylab("normalized alpha diversity") +
    ylim(-1,0) + 
   xlim(50, 110) +
    theme_bw(base_size=40) +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
          plot.title = element_text(hjust = 0.5),
          legend.position = "none") +
  ggtitle("Family", subtitle=paste(
      paste0("b=",est.family), 
      paste0("p=",pval.family)
      )) +
  xlab("age")
    


#For order level
p.o_age <- ggplot(metaphlan.ctds.order, aes(x=c_age_years, y=statistic)) + 
    geom_point() +
  geom_smooth(method="lm") +
    ylim(-1,0) + 
   xlim(50, 110) +
    theme_bw(base_size=40) +
    theme(axis.title.y=element_blank(),
          axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
          plot.title = element_text(hjust = 0.5),
          legend.position = "none") +
    ggtitle("Order", subtitle=paste(
      paste0("b=",est.order), 
      paste0("p=",pval.order)
      )) +
  xlab("age")


#For class level
p.c_age <- ggplot(metaphlan.ctds.class, aes(x=c_age_years, y=statistic)) + 
    geom_point() + 
  geom_smooth(method="lm") +
    ylim(-1,0) + 
   xlim(50, 110) +
    theme_bw(base_size=40) +
    theme(axis.title.y=element_blank(),
          axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
          plot.title = element_text(hjust = 0.5),
          legend.position = "none") +
    ggtitle("Class", subtitle=paste(
      paste0("b=",est.class), 
      paste0("p=",pval.class)
      ))
  xlab("age")


#For phylum level
p.p_age <- ggplot(metaphlan.ctds.phylum, aes(x=c_age_years, y=statistic)) + 
    geom_point() +
  geom_smooth(method="lm") +
  ylab("normalized alpha diversity") +
    ylim(-1,0) + 
   xlim(50, 110) +
    theme_bw(base_size=40) +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
          plot.title = element_text(hjust = 0.5),
          legend.position = "none") +
    ggtitle("Phylum", subtitle=paste(
      paste0("b=",est.phylum), 
      paste0("p=",pval.phylum)
      )) +
  xlab("age")


#Plot together in one panel (30 by 8)
ggpubr::ggarrange(p.p_age, p.c_age, p.o_age, p.f_age, p.g_age, p.s_age, 
          ncol = 6, nrow = 1)

#create directory
dir.create(paste0(outpath,"Figures/incommon_taxa/"))

png(paste0(outpath,"Figures/incommon_taxa/entropy_metaphlan_incommon.taxa_chinese_alllevels.png"), units="in", width=30, height=20, res=300)
ggpubr::ggarrange(p.p_age, p.c_age, p.o_age, p.f_age, p.g_age, p.s_age,
           ncol = 3, nrow = 2)
 dev.off()
 
png(paste0(outpath,"Figures/incommon_taxa/entropy_metaphlan_incommon.taxa_chinese_main.png"), units="in", width=30, height=10, res=300)
ggpubr::ggarrange(p.p_age, p.g_age, p.s_age,
           ncol = 3, nrow = 1)
 dev.off()
```

## Bracken: Calculate alpha diversity using otu_ctds function

```{r, warning=FALSE}
#taxa diversity for species
OTU_species <- bracken_chinese_renorm.otu
bracken.ctds.species <- otu_ctds(taxadata = filtered_taxa_total, otudata = OTU_species, rank = c("kingdom","phylum","class","order","family","genus","species"), sample.meta = bracken.sample.meta)

#taxa diversity for genus
OTU_genus <- bracken_chinese_renorm.otu
bracken.ctds.genus <- otu_ctds(taxadata = filtered_taxa_total, otudata = OTU_genus, rank = c("kingdom","phylum","class","order","family","genus"), sample.meta = bracken.sample.meta)

#taxa diversity for family
OTU_family <- bracken_chinese_renorm.otu
bracken.ctds.family <- otu_ctds(taxadata = filtered_taxa_total, otudata = OTU_family, rank = c("kingdom","phylum","class","order","family"), sample.meta = bracken.sample.meta)

#taxa diversity for order
OTU_order <- bracken_chinese_renorm.otu
bracken.ctds.order <- otu_ctds(taxadata = filtered_taxa_total, otudata = OTU_order, rank = c("kingdom","phylum","class","order"), sample.meta = bracken.sample.meta)

#taxa diversity for class
OTU_class <- bracken_chinese_renorm.otu
bracken.ctds.class <- otu_ctds(taxadata = filtered_taxa_total, otudata = OTU_class, rank = c("kingdom","phylum","class"), sample.meta = bracken.sample.meta)

#taxa diversity for phylum
OTU_phylum <- bracken_chinese_renorm.otu
bracken.ctds.phylum <- otu_ctds(taxadata = filtered_taxa_total, otudata = OTU_phylum, rank = c("kingdom","phylum"), sample.meta = bracken.sample.meta)


#extract scores and save as csv file
bracken.ctds.phylum.score <- bracken.ctds.phylum %>% dplyr::select(sample.ID, statistic) %>% dplyr::rename(phylum_statistic=statistic)
bracken.ctds.class.score <- bracken.ctds.class %>% dplyr::select(statistic) %>% dplyr::rename(class_statistic=statistic)
bracken.ctds.order.score <- bracken.ctds.order %>% dplyr::select(statistic) %>% dplyr::rename(order_statistic=statistic)
bracken.ctds.family.score <- bracken.ctds.family %>% dplyr::select(statistic) %>% dplyr::rename(family_statistic=statistic)
bracken.ctds.genus.score <- bracken.ctds.genus %>% dplyr::select(statistic) %>% dplyr::rename(genus_statistic=statistic)
bracken.ctds.species.score <- bracken.ctds.species %>% dplyr::select(statistic) %>% dplyr::rename(species_statistic=statistic)

bracken.ctds.list <- dplyr::bind_cols(bracken.ctds.phylum.score, bracken.ctds.class.score, bracken.ctds.order.score, bracken.ctds.family.score, bracken.ctds.genus.score, bracken.ctds.species.score)
                  
write_csv(bracken.ctds.list, file=paste0(outpath,"entropy.bracken.commontaxa.chinese.csv"))

```

### Age association

```{r}
#species level
lm.species <- lm(statistic ~ c_age_years, data=bracken.ctds.species)
pval.species <- broom::tidy(lm.species) %>% dplyr::filter(term=="c_age_years") %>% pull(p.value) %>% signif(3)
est.species <- broom::tidy(lm.species) %>% dplyr::filter(term=="c_age_years") %>% pull(estimate) %>% signif(3)
summary(lm.species)

#genus level
lm.genus <- lm(statistic ~ c_age_years, data=bracken.ctds.genus)
pval.genus <- broom::tidy(lm.genus) %>% dplyr::filter(term=="c_age_years") %>% pull(p.value) %>% signif(3)
est.genus <- broom::tidy(lm.genus) %>% dplyr::filter(term=="c_age_years") %>% pull(estimate) %>% signif(3)
summary(lm.genus)

#family level
lm.family <- lm(statistic ~ c_age_years, data=bracken.ctds.family)
pval.family <- broom::tidy(lm.family) %>% dplyr::filter(term=="c_age_years") %>% pull(p.value) %>% signif(3)
est.family <- broom::tidy(lm.family) %>% dplyr::filter(term=="c_age_years") %>% pull(estimate) %>% signif(3)
summary(lm.family)

#order level
lm.order <- lm(statistic ~ c_age_years, data=bracken.ctds.order)
pval.order <- broom::tidy(lm.order) %>% dplyr::filter(term=="c_age_years") %>% pull(p.value) %>% signif(3)
est.order <- broom::tidy(lm.order) %>% dplyr::filter(term=="c_age_years") %>% pull(estimate) %>% signif(3)
summary(lm.order)

#class level
lm.class <- lm(statistic ~ c_age_years, data=bracken.ctds.class)
pval.class <- broom::tidy(lm.class) %>% dplyr::filter(term=="c_age_years") %>% pull(p.value) %>% signif(3)
est.class <- broom::tidy(lm.class) %>% dplyr::filter(term=="c_age_years") %>% pull(estimate) %>% signif(3)
summary(lm.class)

#phylum level
lm.phylum <- lm(statistic ~ c_age_years, data=bracken.ctds.phylum)
pval.phylum <- broom::tidy(lm.phylum) %>% dplyr::filter(term=="c_age_years") %>% pull(p.value) %>% signif(3)
est.phylum <- broom::tidy(lm.phylum) %>% dplyr::filter(term=="c_age_years") %>% pull(estimate) %>% signif(3)
summary(lm.phylum)
```

### Plot Alpha diversity across taxonomic levels

```{r, fig.dim = c(50,10)}
#For species level
p.s_age <- ggplot(bracken.ctds.species, aes(x=c_age_years, y=statistic)) + 
    geom_point() + 
  geom_smooth(method="lm") +
    ylim(-1,0) + 
   xlim(50, 110) +
    theme_bw(base_size=40) +
    theme(axis.title.y=element_blank(),
          axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
          plot.title = element_text(hjust = 0.5),
          legend.position = "none") +
    ggtitle("Species", subtitle=paste(
      paste0("b=",est.species), 
      paste0("p=",pval.species)
      )) +
  xlab("age")
    

#For genus level
p.g_age <- ggplot(bracken.ctds.genus, aes(x=c_age_years, y=statistic)) + 
    geom_point() + 
  geom_smooth(method="lm") +
    ylim(-1,0) +
   xlim(50, 110) +
    theme_bw(base_size=40) +
    theme(axis.title.y=element_blank(),
          axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
          plot.title = element_text(hjust = 0.5),
          legend.position = "none") +
    ggtitle("Genus", subtitle=paste(
      paste0("b=",est.genus), 
      paste0("p=",pval.genus)
      )) +
  xlab("age")
    

#For family level
p.f_age <- ggplot(bracken.ctds.family, aes(x=c_age_years, y=statistic)) + 
    geom_point() + 
  geom_smooth(method="lm") +
  ylab("normalized alpha diversity") +
    ylim(-1,0) + 
   xlim(50, 110) +
    theme_bw(base_size=40) +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
          plot.title = element_text(hjust = 0.5),
          legend.position = "none") +
  ggtitle("Family", subtitle=paste(
      paste0("b=",est.family), 
      paste0("p=",pval.family)
      )) +
  xlab("age")
    


#For order level
p.o_age <- ggplot(bracken.ctds.order, aes(x=c_age_years, y=statistic)) + 
    geom_point() +
  geom_smooth(method="lm") +
    ylim(-1,0) + 
   xlim(50, 110) +
    theme_bw(base_size=40) +
    theme(axis.title.y=element_blank(),
          axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
          plot.title = element_text(hjust = 0.5),
          legend.position = "none") +
    ggtitle("Order", subtitle=paste(
      paste0("b=",est.order), 
      paste0("p=",pval.order)
      )) +
  xlab("age")


#For class level
p.c_age <- ggplot(bracken.ctds.class, aes(x=c_age_years, y=statistic)) + 
    geom_point() + 
  geom_smooth(method="lm") +
    ylim(-1,0) + 
   xlim(50, 110) +
    theme_bw(base_size=40) +
    theme(axis.title.y=element_blank(),
          axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
          plot.title = element_text(hjust = 0.5),
          legend.position = "none") +
    ggtitle("Class", subtitle=paste(
      paste0("b=",est.class), 
      paste0("p=",pval.class)
      ))
  xlab("age")


#For phylum level
p.p_age <- ggplot(bracken.ctds.phylum, aes(x=c_age_years, y=statistic)) + 
    geom_point() +
  geom_smooth(method="lm") +
  ylab("normalized alpha diversity") +
    ylim(-1,0) + 
   xlim(50, 110) +
    theme_bw(base_size=40) +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
          plot.title = element_text(hjust = 0.5),
          legend.position = "none") +
    ggtitle("Phylum", subtitle=paste(
      paste0("b=",est.phylum), 
      paste0("p=",pval.phylum)
      )) +
  xlab("age")


#Plot together in one panel (30 by 8)
ggpubr::ggarrange(p.p_age, p.c_age, p.o_age, p.f_age, p.g_age, p.s_age, 
          ncol = 6, nrow = 1)

png(paste0(outpath,"Figures/incommon_taxa/entropy_bracken_incommon.taxa_chinese_alllevels.png"), units="in", width=30, height=20, res=300)
ggpubr::ggarrange(p.p_age, p.c_age, p.o_age, p.f_age, p.g_age, p.s_age,
           ncol = 3, nrow = 2)
 dev.off()
 
 png(paste0(outpath,"Figures/incommon_taxa/entropy_bracken_incommon.taxa_chinese_main.png"), units="in", width=30, height=10, res=300)
ggpubr::ggarrange(p.p_age, p.g_age, p.s_age,
           ncol = 3, nrow = 1)
 dev.off()
```

# Differences in alpha diversity based on method

```{r}
#extract scores for each method
#bracken
bracken.ctds.phylum.score <- bracken.ctds.phylum %>% dplyr::select(sample.ID, statistic) %>% dplyr::rename(bracken_statistic=statistic)
bracken.ctds.class.score <- bracken.ctds.class %>% dplyr::select(sample.ID, statistic) %>% dplyr::rename(bracken_statistic=statistic)
bracken.ctds.order.score <- bracken.ctds.order %>% dplyr::select(sample.ID, statistic) %>% dplyr::rename(bracken_statistic=statistic)
bracken.ctds.family.score <- bracken.ctds.family %>% dplyr::select(sample.ID, statistic) %>% dplyr::rename(bracken_statistic=statistic)
bracken.ctds.genus.score <- bracken.ctds.genus %>% dplyr::select(sample.ID, statistic) %>% dplyr::rename(bracken_statistic=statistic)
bracken.ctds.species.score <- bracken.ctds.species %>% dplyr::select(sample.ID, statistic) %>% dplyr::rename(bracken_statistic=statistic)

#metaphlan
metaphlan.ctds.phylum.score <- metaphlan.ctds.phylum %>% dplyr::select(sample.ID, statistic) %>% dplyr::rename(metaphlan_statistic=statistic)
metaphlan.ctds.class.score <- metaphlan.ctds.class %>% dplyr::select(sample.ID, statistic) %>% dplyr::rename(metaphlan_statistic=statistic)
metaphlan.ctds.order.score <- metaphlan.ctds.order %>% dplyr::select(sample.ID, statistic) %>% dplyr::rename(metaphlan_statistic=statistic)
metaphlan.ctds.family.score <- metaphlan.ctds.family %>% dplyr::select(sample.ID, statistic) %>% dplyr::rename(metaphlan_statistic=statistic)
metaphlan.ctds.genus.score <- metaphlan.ctds.genus %>% dplyr::select(sample.ID, statistic) %>% dplyr::rename(metaphlan_statistic=statistic)
metaphlan.ctds.species.score <- metaphlan.ctds.species %>% dplyr::select(sample.ID, statistic) %>% dplyr::rename(metaphlan_statistic=statistic)

#combine data
combined.ctds.phylum <- dplyr::inner_join(bracken.ctds.phylum.score, metaphlan.ctds.phylum.score, by="sample.ID")
combined.ctds.class <- dplyr::inner_join(bracken.ctds.class.score, metaphlan.ctds.class.score, by="sample.ID")
combined.ctds.order <- dplyr::inner_join(bracken.ctds.order.score, metaphlan.ctds.order.score, by="sample.ID")
combined.ctds.family <- dplyr::inner_join(bracken.ctds.family.score, metaphlan.ctds.family.score, by="sample.ID")
combined.ctds.genus <- dplyr::inner_join(bracken.ctds.genus.score, metaphlan.ctds.genus.score, by="sample.ID")
combined.ctds.species <- dplyr::inner_join(bracken.ctds.species.score, metaphlan.ctds.species.score, by="sample.ID")



```

### Correlation between adjusted entropy based on method

```{r}

#species level
cor.species <- cor.test(combined.ctds.species$metaphlan_statistic, combined.ctds.species$bracken_statistic)
est.species <- broom::tidy(cor.species) %>% pull(estimate) %>% signif(2)
pval.species <- broom::tidy(cor.species) %>% pull(p.value) %>% signif(3)
cor.species

#genus level
cor.genus <- cor.test(combined.ctds.genus$metaphlan_statistic, combined.ctds.genus$bracken_statistic)
est.genus <- broom::tidy(cor.genus) %>% pull(estimate) %>% signif(2)
pval.genus <- broom::tidy(cor.genus) %>% pull(p.value) %>% signif(3)
cor.genus

#family level
cor.family <- cor.test(combined.ctds.family$metaphlan_statistic, combined.ctds.family$bracken_statistic)
est.family <- broom::tidy(cor.family) %>% pull(estimate) %>% signif(2)
pval.family <- broom::tidy(cor.family) %>% pull(p.value) %>% signif(3)
cor.family

#order level
cor.order <- cor.test(combined.ctds.order$metaphlan_statistic, combined.ctds.order$bracken_statistic)
est.order <- broom::tidy(cor.order) %>% pull(estimate) %>% signif(2)
pval.order <- broom::tidy(cor.order) %>% pull(p.value) %>% signif(3)
cor.order

#class level
cor.class <- cor.test(combined.ctds.class$metaphlan_statistic, combined.ctds.class$bracken_statistic)
est.class <- broom::tidy(cor.class) %>% pull(estimate) %>% signif(2)
pval.class <- broom::tidy(cor.class) %>% pull(p.value) %>% signif(3)
cor.class

#phylum level
cor.phylum <- cor.test(combined.ctds.phylum$metaphlan_statistic, combined.ctds.phylum$bracken_statistic)
est.phylum <- broom::tidy(cor.phylum) %>% pull(estimate) %>% signif(2)
pval.phylum <- broom::tidy(cor.phylum) %>% pull(p.value) %>% signif(3)
cor.phylum

```

### Visualize the differences in adjused entropy based on method

```{r}
#For species level
boxplot(combined.ctds.species$metaphlan_statistic, combined.ctds.species$bracken_statistic, main="Species", xlab="method", ylab="adjusted entropy", ylim=c(-1,0), names=c("MetaPhlAn4", "Bracken"))
    
#For genus level
boxplot(combined.ctds.genus$metaphlan_statistic, combined.ctds.genus$bracken_statistic, main="Genus", xlab="method", ylab="adjusted entropy", ylim=c(-1,0), names=c("MetaPhlAn4", "Bracken"))
    
#For family level
boxplot(combined.ctds.family$metaphlan_statistic, combined.ctds.family$bracken_statistic, main="Family", xlab="method", ylab="adjusted entropy", ylim=c(-1,0), names=c("MetaPhlAn4", "Bracken"))
  
#For order level
boxplot(combined.ctds.order$metaphlan_statistic, combined.ctds.order$bracken_statistic, main="Order", xlab="method", ylab="adjusted entropy", ylim=c(-1,0), names=c("MetaPhlAn4", "Bracken"))
  
#For class level
boxplot(combined.ctds.class$metaphlan_statistic, combined.ctds.class$bracken_statistic, main="Class", xlab="method", ylab="adjusted entropy", ylim=c(-1,0), names=c("MetaPhlAn4", "Bracken"))
  
#For phylum level
boxplot(combined.ctds.phylum$metaphlan_statistic, combined.ctds.phylum$bracken_statistic, main="Phylum", xlab="method", ylab="adjusted entropy", ylim=c(-1,0), names=c("MetaPhlAn4", "Bracken"))
  
```


```{r, fig.dim = c(30,20)}
# Species
# Plot paired scores
p.s_comp <- ggplot(combined.ctds.species, aes(x = metaphlan_statistic, y = bracken_statistic)) +
  geom_point() +
  theme_bw(base_size = 40) +
  xlim(-1,0) +
  ylim(-1,0) +
  geom_abline(intercept = 0 , slope = 1 ) +
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(title = "Species", subtitle = paste0("r=",est.phylum,",p=",pval.phylum), x = "MetaPhlAn4", y = "Bracken")

# genus
# Plot paired scores
p.g_comp <-ggplot(combined.ctds.genus, aes(x = metaphlan_statistic, y = bracken_statistic)) +
  geom_point() +
  theme_bw(base_size = 40) +
  xlim(-1,0) +
  ylim(-1,0) +
  geom_abline(intercept = 0 , slope = 1 ) +
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(title = "Genus", subtitle = paste0("r=",est.genus,",p=",pval.genus), x = "MetaPhlAn4", y = "Bracken")


# family
# Plot paired scores
p.f_comp <-ggplot(combined.ctds.family, aes(x = metaphlan_statistic, y = bracken_statistic)) +
  geom_point() +
  theme_bw(base_size = 40) +
  xlim(-1,0) +
  ylim(-1,0) +
  geom_abline(intercept = 0 , slope = 1 ) +
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(title = "Family", subtitle = paste0("r=",est.family,",p=",pval.family), x = "MetaPhlAn4", y = "Bracken") 


# order
# Plot paired scores
p.o_comp <-ggplot(combined.ctds.order, aes(x = metaphlan_statistic, y = bracken_statistic)) +
  geom_point() +
  theme_bw(base_size = 40) +
  xlim(-1,0) +
  ylim(-1,0) +
  geom_abline(intercept = 0 , slope = 1 ) +
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(title = "Order", subtitle = paste0("r=",est.order,",p=",pval.order), x = "MetaPhlAn4", y = "Bracken") 


# class
# Plot paired scores
p.c_comp <-ggplot(combined.ctds.class, aes(x = metaphlan_statistic, y = bracken_statistic)) +
  geom_point() +
  theme_bw(base_size = 40) +
  xlim(-1,0) +
  ylim(-1,0) +
  geom_abline(intercept = 0 , slope = 1 ) +
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(title = "Class", subtitle = paste0("r=",est.class,",p=",pval.class), x = "MetaPhlAn4", y = "Bracken")


# phylum
# Plot paired scores
p.p_comp <-ggplot(combined.ctds.phylum, aes(x = metaphlan_statistic, y = bracken_statistic)) +
  geom_point() +
  theme_bw(base_size = 40) +
  xlim(-1,0) +
  ylim(-1,0) +
  geom_abline(intercept = 0 , slope = 1 ) + 
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(title = "Phylum", subtitle = paste0("r=",est.phylum,",p=",pval.phylum), x = "MetaPhlAn4", y = "Bracken") 


#Plot together in one panel (30 by 8)
ggpubr::ggarrange(p.p_comp, p.c_comp, p.o_comp, p.f_comp, p.g_comp, p.s_comp,
           ncol = 3, nrow = 2)

#save figures
png(paste0(outpath,"Figures/incommon_taxa/entropy_methodcomp_incommon.taxa_chinese_alllevels.png"), units="in", width=30, height=20, res=300)
ggpubr::ggarrange(p.p_comp, p.c_comp, p.o_comp, p.f_comp, p.g_comp, p.s_comp,
           ncol = 3, nrow = 2)
dev.off()
 
png(paste0(outpath,"Figures/incommon_taxa/entropy_methodcomp_incommon.taxa_chinese_main.png"), units="in", width=30, height=10, res=300)
ggpubr::ggarrange(p.p_comp, p.g_comp, p.s_comp,
           ncol = 3, nrow = 1)
dev.off()

```



