---
title: "Beta diversity analysis with Xu et al chinese cohort data based all taxa between taxonomic classification methods"
author: "Tanya Karagiannis"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    theme: paper
    toc: yes
    toc_collapse: no
    toc_float: yes
  pdf_document:
    toc: yes
---

```{r setup, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
library(dplyr)
library(ggplot2)
library(ggpubr)
library(ape)
library(DESeq2)
library(stringr)
library(reshape2)
library(graphics)

library(codebook)
library(future)
#library(vtable)
library(useful)

library(vegan)
library(data.table)

# Data Cleaning
library(janitor)

# EDA
library(skimr)
library(DataExplorer)

# ggplot2 Helpers
library(ggplot2)
library(ggsignif)
library(scales)
library(phyloseq)
library(tidyverse)
library(tidymodels)
#heatmap
library(pheatmap)
library(pheatmap)
library(ComplexHeatmap)
library(circlize)
library(dendextend)
library(dynamicTreeCut)
#microbiome and single cell specific packages
library(vegan)
library(mia)
library(scater)
#write excel sheets
library(writexl)
library(RColorBrewer)

#specify file path
flpath <- "/restricted/projectnb/uh2-sebas/data/metagenomics/external_cohorts/Han_Chinese_Centenarian/processed_data/data_library/"
outpath <- "/restricted/projectnb/uh2-sebas/analysis/metagenomics/parallel_systems_comparison/02_beta_diversity_analysis/Chinese_cohort/output/"
meta.dir <- "/restricted/projectnb/necs/Data_library/Metabolomics/resources/Xu_plasma_metabolite_analysis/"
sra.dir <- "/restricted/projectnb/uh2-sebas/data/metagenomics/external_cohorts/Han_Chinese_Centenarian/metadata/"

#metaphlan data
metaphlan_otufl <- "metaphlan4_renorm_chinese_phyloseq.10.23.2024.rds"
#bracken data with log10 read count threshold 1 and unique kmer minimizer count threshold 3
bracken_otufl <- "bracken_renorm_chinese_phyloseq.R1.K3.10.22.2024.rds"

#metadata
phenofl <- "phenotype.xlsx"

#set seed
set.seed(157)

```

# Functions

```{r user-defined functions}
## outlier function
is_outlier <- function(x) {
  return(x < quantile(x, 0.25) - 1.5 * IQR(x) | x > quantile(x, 0.75) + 1.5 * IQR(x))
}


# calculate relative abundance/normalized abundance
get_relabund <- function(indat, varname){
  relabund <- as.tibble(indat) %>%
    janitor::adorn_percentages("col") %>%
    column_to_rownames(var = varname) 
}
# 
source("/restricted/projectnb/uh2-sebas/analysis/metagenomics/ye_analyses/Scripts/UserDefinedFunctions/CTDS_function.R")

#function to aggregate otu abundances, calculate proportions and calculated CTDS based on taxonomic hierarchy
otu_ctds <- function(taxadata = biom_chinese_kraken.taxa, 
                     otudata=biom_chinese_kraken.otu, rank=rank, 
                     sample.meta=sample.meta){
  #load ctds function
  source("/restricted/projectnb/uh2-sebas/analysis/metagenomics/ye_analyses/Scripts/UserDefinedFunctions/CTDS_function.R")
  
  #prepare taxonomic and otu data to run diversity analysis
  #convert taxonomic table to a tibble with rownames as a column called tax.ID
  #filter taxonomic table to remove taxa with no species label
  biom_taxdata <- as_tibble(taxadata, rownames = "tax.ID") 
 # message(paste("taxa table rows:",nrow(biom_taxdata)))
  #convert otu table to a tibble with rownames as a column called tax.ID
  biom_otudata <- as_tibble(otudata, rownames = "tax.ID")
  
  #save sample names
  biom_samplenames <- colnames(otudata)
  #message(paste("taxa table rows:",nrow(biom_otudata)))

  
  #taxa dictionary: taxonomy table with taxa ID and taxa names
  taxa_dict <- biom_taxdata %>% 
    dplyr::select(tax.ID, {{rank}}) %>%
    unite("tax.name", {{rank}}, remove = FALSE)

  #use an inner_join function from dplyr to join data from both table based on common tax ID
  #create new variable taxanames to create unique
  biom_join <- dplyr::inner_join(taxa_dict, biom_otudata, by = "tax.ID")
  #message(paste("joined table rows:", nrow(biom_join))) 


  #remove rankings and transpose data for samples in rows and taxa IDs in columns
  #make sure the samples are set as rownames
  # biom_prop <- biom_join %>% 
  #   group_by(tax.name) %>% #use rank groupings
  #   summarise_if(is.numeric, sum) %>% #aggregate OTU abundances for each taxa in the rank
  #   select(c(tax.name, contains(".aggregrated.report"))) %>% #select rank and sample OTU abundances
  #   dplyr::rename(tax.ID = tax.name) %>%
  #   tibble::column_to_rownames(var = "tax.ID") %>% #make tax.ID variable to rownames
  #   t()
  
  biom_prop <- biom_join %>% 
    group_by(tax.name) %>% #use rank groupings
    summarise_if(is.numeric, sum) %>% #aggregate OTU abundances for each taxa in the rank
    select(c(tax.name, biom_samplenames)) %>% #select rank and sample OTU abundances
    column_to_rownames(var = "tax.name") %>% #make tax.name variable to rownames
    t() %>% #transpose data to have taxa in columns and samples as rows
    as_tibble(rownames = "sample.ID") %>% #make transposed data into a tibble with rownames as sample.ID variable
    #janitor::adorn_percentages() %>% #use janitor package to get row proportions for each sample
    column_to_rownames(var = "sample.ID") #add sample IDs back to rownames
  

  #ctds function
  biom.res <- CTDS.score(biom_prop, metadata = sample.meta)
  return(biom.res)
  

}


prop_mat <- function(taxadata = biom_chinese_kraken.taxa, 
                     otudata=biom_chinese_kraken.otu, rank=rank, 
                     sample.meta=sample.meta){

  #convert taxonomic table to a tibble with rownames as a column called tax.ID
  #filter taxonomic table to remove taxa with no taxa classification in any level
  biom_taxdata <- as_tibble(taxadata, rownames = "tax.ID") %>% 
    unite("tax.name", kingdom:{{rank}}, remove = FALSE)

  #row number for taxa data after removing incomplete classification
  message(paste("taxa table rows:",nrow(biom_taxdata)))

  #convert otu table to a tibble with rownames as a column called tax.ID
  biom_otudata <- as_tibble(otudata, rownames = "tax.ID")
  #row number for otu data before removing incomplete classification
  message(paste("otu table rows:",nrow(biom_otudata)))

  #use an inner_join function from dplyr to join data from both table based on common tax ID
  biom_join <- inner_join(biom_taxdata, biom_otudata, by = "tax.ID")
  #row number for joined data
  message(paste("joined table rows:", nrow(biom_join))) #row number 11039 for the common taxa
  
  biom_taxid <- biom_join %>% select(c(tax.ID, tax.name))
  

  #remove rankings and transpose data for samples in rows and taxa IDs in columns
  #make sure the samples are set as rownames
  biom_prop <- biom_join %>% 
    group_by(tax.name) %>%
    summarise_if(is.numeric, sum) %>% 
    select(c(tax.name, contains(".aggregrated.report"))) %>%
    column_to_rownames(var = "tax.name") %>% #make tax.name variable to rownames
    t() %>% #transpose data to have taxa in columns and samples as rows
    as_tibble(rownames = "sample.ID") %>% #make transposed data into a tibble with rownames as sample.ID variable
    janitor::adorn_percentages() %>% #use janitor package to get row proportions for each sample
    column_to_rownames(var = "sample.ID") #add sample IDs back to rownames
  dim(biom_prop)
  useful::corner(biom_prop) #use this to visualize a corner of the data

  #rerun abundance table
  biom.res <- CTDS.score(biom_prop, metadata = sample.meta)
  return(biom.res)

}
```

# Read in phenotypic data


```{r, message=FALSE}
# Extract phenotypic data
#load clinical metadata information
#clinical.data <- read_excel(paste0(meta.dir, "DifferentialAbundance_Supplement.xlsx"), sheet = "Table S1 Metadata", skip = 1)
clinical.data <- read_excel(paste0(meta.dir, phenofl))

#add ethnicity variable
clinical.data$ethnicity <- "han_chinese"


#extract sample IDs and create family ID and group ID
sample.ID <- clinical.data %>% dplyr::select(1) %>% pull()

fam.group <- str_remove(sample.ID, "QCS-") %>%                       
              str_split("-") %>%
                unlist() 

fam.group <- matrix(fam.group, nrow=2)

#create family ID
fam.id <- fam.group[1,]

#create group ID
group.id <- fam.group[2,]
group.id[group.id == "1"] <- "Advanced Age"
group.id[group.id == "2" | group.id == "3"] <- "Offspring"


#add to clinical data
clinical.clean <- clinical.data %>%
  mutate(sample.ID = sample.ID, fam.id = fam.id, group = group.id) 
clinical.clean$"C-age (years)" <- as.integer(clinical.clean$"C-age (years)")


#load sample Run names and sample SRA names
SRA.data <- read_csv(paste0(sra.dir,"SraRunInfo.csv"))

SRA.clean <- SRA.data %>% 
  dplyr::select(Run, SampleName) 

#sample ID and sample SRA names
sample.dict <- read_excel(paste0(sra.dir, "sample name-sequencing ID sheet.xlsx"))

#combine files
sample.info <- dplyr::inner_join(SRA.clean, sample.dict, by = c("SampleName"))

pheno.data <- dplyr::inner_join(sample.info, clinical.clean, by = c("sample.ID")) %>% 
                                  dplyr::select(-c(sample.ID)) %>%
                                  mutate(stool_kit = Run, Sample=Run) %>%
                                  dplyr::rename(c_age_years="C-age (years)")
pheno.data$stool_kit <- as.character(pheno.data$stool_kit)
pheno.data$Sample <- as.character(pheno.data$Sample)


SAMPLE_phy <- sample_data(pheno.data %>% column_to_rownames(var="Sample"))
```


# Metaphlan4: Read in phenotype data and phyloseq object
```{r metaphlan data}
#load phyloseq obj
metaphlan_chinese_biom <- readRDS(paste0(flpath, metaphlan_otufl))

#add phenotypic data to phyloseq object
metaphlan_chinese_biom@sam_data <- SAMPLE_phy

#keep samples with age group label
metaphlan_chinese_biom <- subset_samples(metaphlan_chinese_biom, !is.na(c_age_years))

metaphlan.sample.meta <- sample_data(metaphlan_chinese_biom) %>% 
  as_tibble(rownames="stool_kit")
# metaphlan.samples <- metaphlan.sample.meta %>% dplyr::select(sample.ID, stool_kit)
  
OTU_metaphlan <- metaphlan_chinese_biom@otu_table@.Data
# colnames(OTU_metaphlan) <- metaphlan.samples$stool_kit
TAX_metaphlan <- metaphlan_chinese_biom@tax_table@.Data

#save sample data for permanova
# rownames(metaphlan_chinese_biom@sam_data) <- metaphlan_chinese_biom@sam_data$stool_kit
# sample_names(metaphlan_chinese_biom) <- metaphlan_chinese_biom@sam_data$stool_kit
metaphlan.samp.dat <- data.frame(metaphlan_chinese_biom@sam_data)

```

# Bracken: Read in phenotype data and phyloseq object
```{r bracken data}
#load phyloseq obj
bracken_chinese_biom <- readRDS(paste0(flpath, bracken_otufl))

#add phenotypic data to phyloseq object
bracken_chinese_biom@sam_data <- SAMPLE_phy

#keep samples with age group label
bracken_chinese_biom <- subset_samples(bracken_chinese_biom, !is.na(c_age_years))

bracken.sample.meta <- sample_data(bracken_chinese_biom) %>%
  as_tibble(rownames="stool_kit")
# bracken.samples <- bracken.sample.meta %>% dplyr::select(sample.ID, stool_kit)

#calculate relative abundances
OTU_bracken <- bracken_chinese_biom@otu_table@.Data
TAX_bracken <- bracken_chinese_biom@tax_table@.Data

#save sample data for permanova
# rownames(bracken_chinese_biom@sam_data) <- bracken_chinese_biom@sam_data$stool_kit
# sample_names(bracken_chinese_biom) <- bracken_chinese_biom@sam_data$stool_kit
bracken.samp.dat <- data.frame(bracken_chinese_biom@sam_data)

```

## common samples

```{r}
#metaphlan
metaphlan.sample.meta <- metaphlan.sample.meta %>% dplyr::filter(metaphlan.sample.meta$stool_kit %in% bracken.sample.meta$stool_kit)
dim(metaphlan.sample.meta)

OTU_metaphlan <- OTU_metaphlan[, metaphlan.sample.meta$stool_kit]
dim(OTU_metaphlan)

#bracken
bracken.sample.meta <- bracken.sample.meta %>% dplyr::filter(bracken.sample.meta$stool_kit %in% metaphlan.sample.meta$stool_kit)
dim(bracken.sample.meta)

OTU_bracken <- OTU_bracken[, bracken.sample.meta$stool_kit]
dim(OTU_bracken)

#make sure to have the same order of samples
sample_names <- unique(colnames(OTU_bracken), colnames(OTU_metaphlan))
```

# Beta diversity across taxonomic hierarchy

## Metaphlan: Create Tree Summarized Experiment object from Phyloseq object 

```{r metaphlan tse}
#convert phyloseq object to a tree summarized experiment object
metaphlan_tse <- mia::makeTreeSummarizedExperimentFromPhyloseq(metaphlan_chinese_biom)
metaphlan_tse


#agglomerate based on each taxonomic rank and save as alternative assays in the object
altExps(metaphlan_tse) <- mia::splitByRanks(metaphlan_tse)
metaphlan_tse


```

## Bracken: Create Tree Summarized Experiment object from Phyloseq object 

```{r bracken tse}
#convert phyloseq object to a tree summarized experiment object
bracken_tse <- mia::makeTreeSummarizedExperimentFromPhyloseq(bracken_chinese_biom)
bracken_tse


#agglomerate based on each taxonomic rank and save as alternative assays in the object
altExps(bracken_tse) <- mia::splitByRanks(bracken_tse)
bracken_tse


```

## Metaphlan: calculate bray curtis index and uniqueness

```{r metaphlan BC}
#taxa diversity for species
metaphlan_species_tse <- altExp(metaphlan_tse,"Species")
OTU_species <- assay(metaphlan_species_tse, "counts")[,sample_names]
metaphlan_beta.div.species <- rbiom::beta.div(OTU_species, method="bray-curtis", weighted=TRUE)
metaphlan_uniqueness.species <- apply(as.matrix(metaphlan_beta.div.species), 1, FUN = function(x) {min(x[x > 0])})

#join with metadata
metaphlan_b.species <- metaphlan_uniqueness.species %>% as_tibble(rownames="stool_kit") %>%
  dplyr::rename("uniqueness"=value) %>%
  dplyr::inner_join(metaphlan.sample.meta, by = "stool_kit")

#taxa diversity for genus
metaphlan_genus_tse <- altExp(metaphlan_tse,"Genus")
OTU_genus <- assay(metaphlan_genus_tse, "counts")[,sample_names]
metaphlan_beta.div.genus <- rbiom::beta.div(OTU_genus, method="bray-curtis", weighted=TRUE)
metaphlan_uniqueness.genus <- apply(as.matrix(metaphlan_beta.div.genus), 1, FUN = function(x) {min(x[x > 0])})

#join with metadata
metaphlan_b.genus <- metaphlan_uniqueness.genus %>% as_tibble(rownames="stool_kit") %>%
  dplyr::rename("uniqueness"=value) %>%
  dplyr::inner_join(metaphlan.sample.meta, by = "stool_kit")

#taxa diversity for family
metaphlan_family_tse <- altExp(metaphlan_tse,"Family")
OTU_family <- assay(metaphlan_family_tse, "counts")[,sample_names]
metaphlan_beta.div.family <- rbiom::beta.div(OTU_family, method="bray-curtis", weighted=TRUE)
metaphlan_uniqueness.family <- apply(as.matrix(metaphlan_beta.div.family), 1, FUN = function(x) {min(x[x > 0])})

#join with metadata
metaphlan_b.family <- metaphlan_uniqueness.family %>% as_tibble(rownames="stool_kit") %>%
  dplyr::rename("uniqueness"=value) %>%
  dplyr::inner_join(metaphlan.sample.meta, by = "stool_kit")

#taxa diversity for order
metaphlan_order_tse <- altExp(metaphlan_tse,"Order")
OTU_order <- assay(metaphlan_order_tse, "counts")[,sample_names]
metaphlan_beta.div.order <- rbiom::beta.div(OTU_order, method="bray-curtis", weighted=TRUE)
metaphlan_uniqueness.order <- apply(as.matrix(metaphlan_beta.div.order), 1, FUN = function(x) {min(x[x > 0])})

#join with metadata
metaphlan_b.order <- metaphlan_uniqueness.order %>% as_tibble(rownames="stool_kit") %>%
  dplyr::rename("uniqueness"=value) %>%
  dplyr::inner_join(metaphlan.sample.meta, by = "stool_kit")

#taxa diversity for class
metaphlan_class_tse <- altExp(metaphlan_tse,"Class")
OTU_class <- assay(metaphlan_class_tse, "counts")[,sample_names]
metaphlan_beta.div.class <- rbiom::beta.div(OTU_class, method="bray-curtis", weighted=TRUE)
metaphlan_uniqueness.class <- apply(as.matrix(metaphlan_beta.div.class), 1, FUN = function(x) {min(x[x > 0])})

#join with metadata
metaphlan_b.class <- metaphlan_uniqueness.class %>% as_tibble(rownames="stool_kit") %>%
  dplyr::rename("uniqueness"=value) %>%
  dplyr::inner_join(metaphlan.sample.meta, by = "stool_kit")

#taxa diversity for phylum
metaphlan_phylum_tse <- altExp(metaphlan_tse,"Phylum")
OTU_phylum <- assay(metaphlan_phylum_tse, "counts")[,sample_names]
metaphlan_beta.div.phylum <- rbiom::beta.div(OTU_phylum, method="bray-curtis", weighted=TRUE)
metaphlan_uniqueness.phylum <- apply(as.matrix(metaphlan_beta.div.phylum), 1, FUN = function(x) {min(x[x > 0])})

#join with metadata
metaphlan_b.phylum <- metaphlan_uniqueness.phylum %>% as_tibble(rownames="stool_kit") %>%
  dplyr::rename("uniqueness"=value) %>%
  dplyr::inner_join(metaphlan.sample.meta, by = "stool_kit")

```

## Bracken: calculate bray curtis index and uniqueness

```{r bracken BC}
#taxa diversity for species
bracken_species_tse <- altExp(bracken_tse,"species")
OTU_species <- assay(bracken_species_tse, "counts")[,sample_names]
bracken_beta.div.species <- rbiom::beta.div(OTU_species, method="bray-curtis", weighted=TRUE)
bracken_uniqueness.species <- apply(as.matrix(bracken_beta.div.species), 1, FUN = function(x) {min(x[x > 0])})

#join with metadata
bracken_b.species <- bracken_uniqueness.species %>% as_tibble(rownames="stool_kit") %>%
  dplyr::rename("uniqueness"=value) %>%
  dplyr::inner_join(bracken.sample.meta, by = "stool_kit")

#taxa diversity for genus
bracken_genus_tse <- altExp(bracken_tse,"genus")
OTU_genus <- assay(bracken_genus_tse, "counts")[,sample_names]
bracken_beta.div.genus <- rbiom::beta.div(OTU_genus, method="bray-curtis", weighted=TRUE)
bracken_uniqueness.genus <- apply(as.matrix(bracken_beta.div.genus), 1, FUN = function(x) {min(x[x > 0])})

#join with metadata
bracken_b.genus <- bracken_uniqueness.genus %>% as_tibble(rownames="stool_kit") %>%
  dplyr::rename("uniqueness"=value) %>%
  dplyr::inner_join(bracken.sample.meta, by = "stool_kit")

#taxa diversity for family
bracken_family_tse <- altExp(bracken_tse,"family")
OTU_family <- assay(bracken_family_tse, "counts")[,sample_names]
bracken_beta.div.family <- rbiom::beta.div(OTU_family, method="bray-curtis", weighted=TRUE)
bracken_uniqueness.family <- apply(as.matrix(bracken_beta.div.family), 1, FUN = function(x) {min(x[x > 0])})

#join with metadata
bracken_b.family <- bracken_uniqueness.family %>% as_tibble(rownames="stool_kit") %>%
  dplyr::rename("uniqueness"=value) %>%
  dplyr::inner_join(bracken.sample.meta, by = "stool_kit")

#taxa diversity for order
bracken_order_tse <- altExp(bracken_tse,"order")
OTU_order <- assay(bracken_order_tse, "counts")[,sample_names]
bracken_beta.div.order <- rbiom::beta.div(OTU_order, method="bray-curtis", weighted=TRUE)
bracken_uniqueness.order <- apply(as.matrix(bracken_beta.div.order), 1, FUN = function(x) {min(x[x > 0])})

#join with metadata
bracken_b.order <- bracken_uniqueness.order %>% as_tibble(rownames="stool_kit") %>%
  dplyr::rename("uniqueness"=value) %>%
  dplyr::inner_join(bracken.sample.meta, by = "stool_kit")

#taxa diversity for class
bracken_class_tse <- altExp(bracken_tse,"class")
OTU_class <- assay(bracken_class_tse, "counts")[,sample_names]
bracken_beta.div.class <- rbiom::beta.div(OTU_class, method="bray-curtis", weighted=TRUE)
bracken_uniqueness.class <- apply(as.matrix(bracken_beta.div.class), 1, FUN = function(x) {min(x[x > 0])})

#join with metadata
bracken_b.class <- bracken_uniqueness.class %>% as_tibble(rownames="stool_kit") %>%
  dplyr::rename("uniqueness"=value) %>%
  dplyr::inner_join(bracken.sample.meta, by = "stool_kit")

#taxa diversity for phylum
bracken_phylum_tse <- altExp(bracken_tse,"phylum")
OTU_phylum <- assay(bracken_phylum_tse, "counts")[,sample_names]
bracken_beta.div.phylum <- rbiom::beta.div(OTU_phylum, method="bray-curtis", weighted=TRUE)
bracken_uniqueness.phylum <- apply(as.matrix(bracken_beta.div.phylum), 1, FUN = function(x) {min(x[x > 0])})

#join with metadata
bracken_b.phylum <- bracken_uniqueness.phylum %>% as_tibble(rownames="stool_kit") %>%
  dplyr::rename("uniqueness"=value) %>%
  dplyr::inner_join(bracken.sample.meta, by = "stool_kit")
```


# Species

## Metaphlan

### PERMANOVA: Association with age

```{r}
samp.dat <- as.matrix(metaphlan.samp.dat)
#species level
metaphlan_dist_matrix.species <- as.matrix(metaphlan_beta.div.species)
samp.dat.species <- as.data.frame(samp.dat[rownames(metaphlan_dist_matrix.species),])

pm.species<-adonis2(metaphlan_dist_matrix.species~as.integer(samp.dat.species$c_age_years), data=samp.dat.species, permutations=999, by="margin")
pm.species
```


## Bracken

### PERMANOVA: Association with age

```{r}
samp.dat <- as.matrix(bracken.samp.dat)
#species level
bracken_dist_matrix.species <- as.matrix(bracken_beta.div.species)
samp.dat.species <- as.data.frame(samp.dat[rownames(bracken_dist_matrix.species),])

pm.species<-adonis2(bracken_dist_matrix.species~as.integer(samp.dat.species$c_age_years), data=samp.dat.species, permutations=999, by="margin")
pm.species
```

### Mantel Test: test how similar the distance matrices are
```{r}
mantel.species <- mantel(metaphlan_beta.div.species, bracken_beta.div.species, method="pearson", permutations=999)
mantel.species
```


# Genus

## Metaphlan

### PERMANOVA: Association with age

```{r}
samp.dat <- as.matrix(metaphlan.samp.dat)
#genus level
metaphlan_dist_matrix.genus <- as.matrix(metaphlan_beta.div.genus)
samp.dat.genus <- as.data.frame(samp.dat[rownames(metaphlan_dist_matrix.genus),])

pm.genus<-adonis2(metaphlan_dist_matrix.genus~as.integer(samp.dat.genus$c_age_years), data=samp.dat.genus, permutations=999, by="margin")
pm.genus
```


## Bracken

### PERMANOVA: Association with age

```{r}
samp.dat <- as.matrix(bracken.samp.dat)
#genus level
bracken_dist_matrix.genus <- as.matrix(bracken_beta.div.genus)
samp.dat.genus <- as.data.frame(samp.dat[rownames(bracken_dist_matrix.genus),])

pm.genus<-adonis2(bracken_dist_matrix.genus~as.integer(samp.dat.genus$c_age_years), data=samp.dat.genus, permutations=999, by="margin")
pm.genus
```

## Mantel Test: test how similar the distance matrices are
```{r}
mantel.genus <- mantel(metaphlan_beta.div.genus, bracken_beta.div.genus, method="pearson", permutations=999)
mantel.genus
```

# Family

## Metaphlan

### PERMANOVA: Association with age

```{r}
samp.dat <- as.matrix(metaphlan.samp.dat)
#family level
metaphlan_dist_matrix.family <- as.matrix(metaphlan_beta.div.family)
samp.dat.family <- as.data.frame(samp.dat[rownames(metaphlan_dist_matrix.family),])

pm.family<-adonis2(metaphlan_dist_matrix.family~as.integer(samp.dat.family$c_age_years), data=samp.dat.family, permutations=999, by="margin")
pm.family
```


## Bracken

### PERMANOVA: Association with age

```{r}
samp.dat <- as.matrix(bracken.samp.dat)
#family level
bracken_dist_matrix.family <- as.matrix(bracken_beta.div.family)
samp.dat.family <- as.data.frame(samp.dat[rownames(bracken_dist_matrix.family),])

pm.family<-adonis2(bracken_dist_matrix.family~as.integer(samp.dat.family$c_age_years), data=samp.dat.family, permutations=999, by="margin")
pm.family
```

## Mantel Test: test how similar the distance matrices are
```{r}
mantel.family <- mantel(metaphlan_beta.div.family, bracken_beta.div.family, method="pearson", permutations=999)
mantel.family
```

# Order

## Metaphlan

### PERMANOVA: Association with age

```{r}
samp.dat <- as.matrix(metaphlan.samp.dat)
#order level
metaphlan_dist_matrix.order <- as.matrix(metaphlan_beta.div.order)
samp.dat.order <- as.data.frame(samp.dat[rownames(metaphlan_dist_matrix.order),])

pm.order<-adonis2(metaphlan_dist_matrix.order~as.integer(samp.dat.order$c_age_years), data=samp.dat.order, permutations=999, by="margin")
pm.order
```


## Bracken

### PERMANOVA: Association with age

```{r}
samp.dat <- as.matrix(bracken.samp.dat)
#order level
bracken_dist_matrix.order <- as.matrix(bracken_beta.div.order)
samp.dat.order <- as.data.frame(samp.dat[rownames(bracken_dist_matrix.order),])

pm.order<-adonis2(bracken_dist_matrix.order~as.integer(samp.dat.order$c_age_years), data=samp.dat.order, permutations=999, by="margin")
pm.order
```

## Mantel Test: test how similar the distance matrices are
```{r}
mantel.order <- mantel(metaphlan_beta.div.order, bracken_beta.div.order, method="pearson", permutations=999)
mantel.order
```

# Class

## Metaphlan

### PERMANOVA: Association with age

```{r}
samp.dat <- as.matrix(metaphlan.samp.dat)
#class level
metaphlan_dist_matrix.class <- as.matrix(metaphlan_beta.div.class)
samp.dat.class <- as.data.frame(samp.dat[rownames(metaphlan_dist_matrix.class),])

pm.class<-adonis2(metaphlan_dist_matrix.class~as.integer(samp.dat.class$c_age_years), data=samp.dat.class, permutations=999, by="margin")
pm.class
```


## Bracken

### PERMANOVA: Association with age

```{r}
samp.dat <- as.matrix(bracken.samp.dat)
#class level
bracken_dist_matrix.class <- as.matrix(bracken_beta.div.class)
samp.dat.class <- as.data.frame(samp.dat[rownames(bracken_dist_matrix.class),])

pm.class<-adonis2(bracken_dist_matrix.class~as.integer(samp.dat.class$c_age_years), data=samp.dat.class, permutations=999, by="margin")
pm.class
```

## Mantel Test: test how similar the distance matrices are
```{r}
mantel.class <- mantel(metaphlan_beta.div.class, bracken_beta.div.class, method="pearson", permutations=999)
mantel.class
```

# Phylum

## Metaphlan

### PERMANOVA: Association with age

```{r}
samp.dat <- as.matrix(metaphlan.samp.dat)
#phylum level
metaphlan_dist_matrix.phylum <- as.matrix(metaphlan_beta.div.phylum)
samp.dat.phylum <- as.data.frame(samp.dat[rownames(metaphlan_dist_matrix.phylum),])

pm.phylum<-adonis2(metaphlan_dist_matrix.phylum~as.integer(samp.dat.phylum$c_age_years), data=samp.dat.phylum, permutations=999, by="margin")
pm.phylum
```


## Bracken

### PERMANOVA: Association with age

```{r}
samp.dat <- as.matrix(bracken.samp.dat)
#phylum level
bracken_dist_matrix.phylum <- as.matrix(bracken_beta.div.phylum)
samp.dat.phylum <- as.data.frame(samp.dat[rownames(bracken_dist_matrix.phylum),])

pm.phylum<-adonis2(bracken_dist_matrix.phylum~as.integer(samp.dat.phylum$c_age_years), data=samp.dat.phylum, permutations=999, by="margin")
pm.phylum
```

## Mantel Test: test how similar the distance matrices are
```{r}
mantel.phylum <- mantel(metaphlan_beta.div.phylum, bracken_beta.div.phylum, method = "pearson", permutations = 999)
mantel.phylum
```

# Visualize Mantel test results

## For all taxonomic levels

```{r, fig.dim=c(25,5)}

# create tibble with all results
mantel.res <- tibble(
    taxonomic_level = factor(c('Phylum', 'Class', 'Order', 'Family', 'Genus', 'Species'), levels=c('Phylum', 'Class', 'Order', 'Family', 'Genus', 'Species')),
    correlation = c(mantel.phylum$statistic, 
                    mantel.class$statistic, 
                    mantel.order$statistic, 
                    mantel.family$statistic, 
                    mantel.genus$statistic, 
                    mantel.species$statistic),
    p_value = c(mantel.phylum$signif,
                mantel.class$signif, 
                mantel.order$signif,
                mantel.family$signif, 
                mantel.genus$signif, 
                mantel.species$signif)
)

# Create the scatter plot
p.mantel <- ggplot(mantel.res, aes(x = taxonomic_level, y = correlation)) +
    geom_bar(stat = "identity", width=0.5) + 
  ylim(0,1) +
    labs(title = "Overall profile comparison: MetaPhlAn4 vs. Bracken",
         y = "correlation coefficient",
         x = "taxonomic level") +
  geom_signif(y_position = c(0.95, 0.95, 0.95, 0.95, 0.95, 0.95), 
              xmin = c(0.8, 1.8, 2.8, 3.8, 4.8, 5.8), 
              xmax = c(1.2, 2.2, 3.2, 4.2, 5.2, 6.2),
              annotation=paste0("p=",mantel.res$p_value), 
              tip_length=0) +
    theme_bw(base_size = 20) +
  theme(axis.title.x=element_blank())

p.mantel

#create directory
dir.create(paste0(outpath,"Figures/all_taxa_method_comparison"))

#save figure
png(paste0(outpath,"Figures/all_taxa_method_comparison/mantel_test_all.taxa_chinese_alllevels.png"), units="in", width=20, height=5, res=300)
p.mantel
dev.off()
```

## For phylum, genus, species level 

```{r, fig.dim=c(15,5)}

# create tibble with all results
mantel.res <- tibble(
    taxonomic_level = factor(c('Phylum', 'Genus', 'Species'), levels=c('Phylum', 'Genus', 'Species')),
    correlation = c(mantel.phylum$statistic, 
                    mantel.genus$statistic, 
                    mantel.species$statistic),
    p_value = c(mantel.phylum$signif,
                mantel.genus$signif, 
                mantel.species$signif)
)

# Create the scatter plot
p.mantel <- ggplot(mantel.res, aes(x = taxonomic_level, y = correlation)) +
    geom_bar(stat = "identity", width=0.5) + 
  ylim(0,1) +
    labs(title = "Overall profile comparison: MetaPhlAn4 vs. Bracken",
         y = "correlation coefficient",
         x = "taxonomic level") +
  geom_signif(y_position = c(0.95, 0.95, 0.95), 
              xmin = c(0.8, 1.8, 2.8), 
              xmax = c(1.2, 2.2, 3.2),
              annotation=paste0("p=",mantel.res$p_value), 
              tip_length=0) +
    theme_bw(base_size = 20) +
  theme(axis.title.x=element_blank())

p.mantel

png(paste0(outpath,"Figures/all_taxa_method_comparison/mantel_test_all.taxa_chinese_main.png"), units="in", width=10, height=5, res=300)
p.mantel
dev.off()
```

